---
title: "Data Overview"
subtitle: "Redd Counts and Environmental Covariates by Core Area and Population"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: right
    toc-title: "Core Areas & Populations"
---

<!-- =====================================================================
     COLLABORATORS: Feel free to edit the introductory text below.
     Please DO NOT modify anything inside ```{r} ... ``` code blocks —
     those generate all the plots and tables automatically.
     ===================================================================== -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r load, include=FALSE}
library(here)
library(tidyverse)
library(patchwork)

source(here("Code", "MPVA_Functions1_13.R"))
suppressMessages(source(here("Code", "PrepareData_MPVA1_9.R")))

# ── Index helpers ─────────────────────────────────────────────────────────────
pop_core_vec <- CoreArea_Num[match(seq_len(nPop), LocalPops_Num)]

# ── Population display names ("Boulder Creek (09040001BLD)") ─────────────────
# PatchWB_In format: "BLTPatchID.WaterbodyName" — one row per survey site
wb_raw <- sapply(PatchWB_In, function(x) strsplit(x, "\\.")[[1]][2])
# Insert spaces before capital letters to recover readable name
wb_readable <- trimws(gsub("([a-z0-9'])([A-Z])", "\\1 \\2", wb_raw))
# Map to populations: use the first site's name for each population
pop_display <- sapply(seq_len(nPop), function(p) {
  first_site <- which(LocalPops_Num == p)[1]
  paste0(wb_readable[first_site], " (", U_LocalPops_In[p], ")")
})
# Anchor-safe ID for each population (lowercase, hyphens)
pop_anchor <- tolower(gsub("[^a-zA-Z0-9]+", "-", U_LocalPops_In))

# ── Core area anchor IDs ──────────────────────────────────────────────────────
core_anchor <- tolower(gsub("[^a-zA-Z0-9]+", "-", U_CoreAreas_In))

# ── Tidy redd counts ──────────────────────────────────────────────────────────
df_redds <- as.data.frame(n_in) |>
  setNames(as.character(DatYears)) |>
  mutate(
    site      = row_number(),
    pop       = LocalPops_Num,
    pop_name  = U_LocalPops_In[LocalPops_Num],
    core      = pop_core_vec[LocalPops_Num],
    core_name = U_CoreAreas_In[core]
  ) |>
  pivot_longer(as.character(DatYears), names_to = "year", values_to = "redds") |>
  mutate(year = as.integer(year))

# Population total per year (sum across sites)
df_pop_total <- df_redds |>
  group_by(pop, pop_name, core, core_name, year) |>
  summarise(total   = sum(redds, na.rm = TRUE),
            any_obs = any(!is.na(redds)),
            .groups = "drop") |>
  mutate(total = if_else(any_obs, total, NA_real_))

# ── Tidy flow (log summer Q, pre-z-score, actual calendar years) ───────────
# xSumFlow is an xtabs-derived matrix; use as.data.frame.matrix() not as.data.frame()
df_flow <- as.data.frame.matrix(xSumFlow) |>
  rownames_to_column("pop_name") |>
  pivot_longer(-pop_name, names_to = "year", values_to = "log_flow") |>
  mutate(year = as.integer(year)) |>
  filter(!is.na(log_flow), is.finite(log_flow))

# ── Tidy temperature (°C, pre-z-score, actual calendar years) ────────────
df_temp <- as.data.frame.matrix(xtTemp) |>
  rownames_to_column("pop_name") |>
  pivot_longer(-pop_name, names_to = "year", values_to = "aug_temp_C") |>
  mutate(year = as.integer(year),
         aug_temp_C = if_else(aug_temp_C == 0, NA_real_, aug_temp_C)) |>
  filter(!is.na(aug_temp_C))

# ── Tidy invasives (raw proportions — do NOT use z-scored logit here) ────
# InvRatIn_raw is z-scored logit; plogis(z-score) is NOT a proportion.
# Read the actual interpolated proportions (0–1) directly.
InvProp_raw <- read.csv(here('Data', 'InvasiveProportion_Proportion_Interpolated.csv'),
                        row.names = 1, check.names = FALSE)
colnames(InvProp_raw) <- gsub("^X", "", colnames(InvProp_raw))

df_inv <- InvProp_raw |>
  as.data.frame() |>
  rownames_to_column("pop_name") |>
  pivot_longer(-pop_name, names_to = "year", values_to = "inv_prop") |>
  mutate(year = as.integer(year),
         # 1e-6 is the numerical floor used to allow logit-transform of 0s;
         # values below 1e-4 are indistinguishable from zero — clip to 0.
         inv_prop = if_else(inv_prop < 1e-4, 0, inv_prop)) |>
  filter(!is.na(inv_prop))

# ── Tidy LKT (core-level; show z-scored logit index) ────────────────────
df_lkt <- LKT_raw |>
  as.data.frame() |>
  rownames_to_column("core_name") |>
  pivot_longer(-core_name, names_to = "year", values_to = "lkt_index") |>
  mutate(year = as.integer(year)) |>
  filter(!is.na(lkt_index))

# ── Static covariate table ────────────────────────────────────────────────
df_static <- tibble(
  pop_name    = U_LocalPops_In,
  pop_display = pop_display,
  core_name   = U_CoreAreas_In[pop_core_vec],
  adfluvial   = AdfluvialIn,
  vb_zscored  = round(VBIn,    2),
  flood_zscored = round(FloodIn, 2),
  rdvb_zscored  = round(RDVBIn,  2),
  extent_km   = round(extent_pop / 1000, 2)
)

# ── Plot helpers ──────────────────────────────────────────────────────────────
theme_data <- theme_bw(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_text(size = 10))

plot_redds <- function(pop_idx) {
  pn <- U_LocalPops_In[pop_idx]
  d_sites  <- df_redds    |> filter(pop == pop_idx, !is.na(redds))
  d_totals <- df_pop_total |> filter(pop == pop_idx, !is.na(total))
  ggplot() +
    geom_col(data = d_totals, aes(x = year, y = total),
             fill = "steelblue", alpha = 0.25, width = 0.8) +
    geom_point(data = d_sites, aes(x = year, y = redds),
               color = "grey30", alpha = 0.7, size = 1.8) +
    geom_line(data = d_totals, aes(x = year, y = total),
              color = "steelblue", linewidth = 0.9) +
    labs(x = NULL, y = "Redd count", title = "Redd counts") +
    theme_data
}

plot_flow <- function(pop_idx) {
  pn <- U_LocalPops_In[pop_idx]
  d <- df_flow |> filter(pop_name == pn)
  if (nrow(d) == 0) return(ggplot() + labs(title = "Flow: no data") + theme_data)
  ggplot(d, aes(x = year, y = log_flow)) +
    geom_line(color = "steelblue3", linewidth = 0.9) +
    geom_point(size = 1.5, color = "steelblue3") +
    labs(x = NULL, y = "Log summer discharge", title = "Summer streamflow") +
    theme_data
}

plot_temp <- function(pop_idx) {
  pn <- U_LocalPops_In[pop_idx]
  d <- df_temp |> filter(pop_name == pn)
  if (nrow(d) == 0) return(ggplot() + labs(title = "Temperature: no data") + theme_data)
  ggplot(d, aes(x = year, y = aug_temp_C)) +
    geom_line(color = "tomato3", linewidth = 0.9) +
    geom_point(size = 1.5, color = "tomato3") +
    labs(x = NULL, y = "August temperature (°C)", title = "August water temperature") +
    theme_data
}

plot_inv <- function(pop_idx) {
  pn <- U_LocalPops_In[pop_idx]
  d <- df_inv |> filter(pop_name == pn)
  if (nrow(d) == 0) return(ggplot() + labs(title = "Invasives: no data") + theme_data)
  ggplot(d, aes(x = year, y = inv_prop)) +
    geom_area(fill = "darkorange", alpha = 0.2) +
    geom_line(color = "darkorange", linewidth = 0.9) +
    geom_point(size = 1.5, color = "darkorange") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       limits = c(0, max(max(d$inv_prop, na.rm = TRUE), 0.01))) +
    labs(x = NULL, y = "Non-native proportion", title = "Non-native trout ratio") +
    theme_data
}

plot_lkt <- function(core_idx) {
  cn <- U_CoreAreas_In[core_idx]
  d <- df_lkt |> filter(core_name == cn)
  if (nrow(d) == 0) return(ggplot() + labs(title = "Lake trout: no data") + theme_data)
  ggplot(d, aes(x = year, y = lkt_index)) +
    geom_hline(yintercept = 0, linetype = 2, color = "grey60") +
    geom_area(fill = "purple4", alpha = 0.15) +
    geom_line(color = "purple4", linewidth = 0.9) +
    geom_point(size = 1.5, color = "purple4") +
    labs(x = NULL, y = "Lake trout index\n(standardized logit)",
         title = "Lake trout abundance index") +
    theme_data
}
```

## Data Summary

This page documents the data assembled across **`r nPop` local populations** in
**`r nCore` core areas**, spanning `r min(DatYears)`–`r max(DatYears)`. All covariates
were compiled from multiple state, federal, and academic sources and processed into
a common format for the hierarchical model. The sections below describe each data
type, its source, and the transformations applied before model fitting.

---

### Population Abundance — Redd Counts

Annual redd counts are the primary abundance index. Spawning redds (nests) are
counted by technicians walking fixed survey reaches each fall; counts are
proportional to the number of spawning adults and have been used to track bull
trout population trends across the recovery units for more than two decades.

**Source:** USGS-harmonized redd count database compiled from USFWS, Montana FWP,
Idaho IDFG, and tribal monitoring programs.

**Coverage:** Survey sites with at least 10 years of data are retained. Within each
population, multiple survey reaches (sites) contribute counts in the same year;
the population-level total is the sum across sites. A count of zero is recorded
when a site was surveyed but no redds were observed; years when a site was not
surveyed are treated as missing (not zero).

**In the model:** Redd counts enter as the observed response. Modeled as NB2
(negative binomial, fixed dispersion) conditioned on latent adult abundance,
survey-reach length, and a contemporaneous flow effect on detection probability.

---

### Dynamic Covariates — Annual Time Series

The following covariates vary by population and year and enter the population
process equation with a **6-year lag** (5 years specified in data prep; the C++
model adds 1 additional year). This lag reflects the approximate time from
egg incubation under a given year's environmental conditions to first-time
spawning as an adult — roughly one bull trout generation.

```{r covar-table, echo=FALSE}
cov_tab <- tibble::tribble(
  ~Covariate,         ~`Spatial scale`, ~`Metric`,                        ~`Transformation`,              ~`Source`,
  "Summer streamflow", "Population",    "Annual sum of daily discharge",  "Log → z-score → 6-yr lag",    "USGS stream gauge network; ProcessedFlowMetrics.csv",
  "August temperature","Population",    "Mean August stream temperature (°C)", "Z-score → 6-yr lag",    "NorWeST spatially-continuous stream temperature model (Isaak et al. 2016); MeanPatchAnnualTemps.RDS",
  "Non-native fish proportion (INV)", "Population", "Proportion of non-native salmonids (brook + brown trout) in electrofishing catch", "Floor at 0.1% → logit → z-score; forward-filled for missing years", "Electrofishing surveys compiled from MFWP, USFWS, Tribes; BLTPatch_InvasiveRatio",
  "Lake trout index (LKT)", "Core area",  "Proportion of lake trout in netting catch", "Logit → z-score; forward-filled for missing years", "Annual netting surveys — Glacier NP and MFWP; LakeNetting/"
)

knitr::kable(cov_tab, format = "html") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"),
                            font_size = 12, full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::column_spec(5, width = "28%")
```

A **contemporaneous flow** term (same year, no lag) is also included in the
*observation* model to account for year-to-year variation in redd detectability —
high-flow years reduce observer efficiency and can suppress counts independently
of true abundance changes.

**Lake trout** data are measured at the **core area** level (not population) and
are only available for lake-connected systems. A binary indicator flags which
core areas have LKT data; the LKT effect is estimated only for those cores.

---

### Static Covariates — Population Characteristics

The following covariates are fixed through time and characterize the physical
setting and life-history context of each population. They inform the
population-level mean growth rate (b₀ᵣ) through a regression on static
population attributes.

```{r static-table, echo=FALSE}
static_tab <- tibble::tribble(
  ~Covariate,                ~`Metric`,                                       ~`Transformation`,          ~`Source`,
  "Valley bottom extent (VB)", "Riparian valley floor area (km²)",           "Log(x + 1) → z-score",    "GIS — Whited, Flathead Lake Biological Station",
  "W95 flood frequency",       "Mean annual days exceeding 95th-pct flow",   "Log(x + 1) → z-score",    "Derived from USGS stream gauge flow records",
  "Road density — valley bottom (RDVB)", "Road length per km² of valley bottom", "Log(x + 1) → z-score", "National road layer intersected with valley bottom GIS — Whited, FLBS",
  "Adfluvial life history",    "Binary: population has lake-resident life-history access", "0/1",         "Core area life-history classification — USFWS Recovery Plan; CoreArea_LifeHistory.csv",
  "Habitat extent (km)",       "Connected spawning habitat length (km)",     "Used as density denominator in process model", "GIS — connected habitat layer, BLTPatches_metrics"
)

knitr::kable(static_tab, format = "html") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "condensed"),
                            font_size = 12, full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::column_spec(4, width = "28%")
```

Valley bottom extent and road density were derived from GIS layers developed by
**Diane Whited** at the Flathead Lake Biological Station. These capture landscape-scale
habitat availability and disturbance pressure that can affect spawning habitat
quality and population productivity independently of year-to-year environmental
variation.

---

### Population Time Series Plots

The plots below show, for each local population: annual redd counts (bars = totals,
points = individual survey-site counts), summer streamflow, August temperature,
and non-native trout proportion. Core area sections also include a lake trout
abundance index (where available) and a static covariate summary table.
Covariate time series are displayed in their native units; the model uses
z-scored and lagged versions.

```{r core-index, results='asis'}
# Clickable core area index
cat("## Jump to core area\n\n")
for (ca in seq_len(nCore)) {
  n_pops <- sum(pop_core_vec == ca)
  cat(sprintf("- [%s](#%s) — %d population%s\n",
              U_CoreAreas_In[ca], core_anchor[ca],
              n_pops, if (n_pops > 1) "s" else ""))
}
cat("\n---\n\n")
```

```{r all-sections, results='asis', fig.width=11, fig.height=9}
for (ca in seq_len(nCore)) {

  core_label   <- U_CoreAreas_In[ca]
  pops_in_core <- which(pop_core_vec == ca)
  n_pops       <- length(pops_in_core)

  # ── Core area header (explicit anchor ID for reliable TOC linking) ──────────
  cat(sprintf("\n\n## %s {#%s}\n\n", core_label, core_anchor[ca]))

  # Summary line
  redd_core  <- df_redds |> filter(core == ca)
  yr_range   <- range(redd_core$year[!is.na(redd_core$redds)])
  cat(sprintf("**%d local population%s** | **%d – %d**\n\n",
              n_pops, if (n_pops > 1) "s" else "", yr_range[1], yr_range[2]))

  # Static covariate table for this core (use display names)
  stab <- df_static |>
    filter(core_name == core_label) |>
    select(Population = pop_display,
           `Adfluvial` = adfluvial,
           `Valley bottom (z)` = vb_zscored,
           `W95 flood frequency (z)` = flood_zscored,
           `Road density VB (z)` = rdvb_zscored,
           `Extent (km)` = extent_km)

  print(
    knitr::kable(stab, format = "html", align = c("l","c","r","r","r","r")) |>
    kableExtra::kable_styling(bootstrap_options = c("striped","condensed"),
                              font_size = 12, full_width = FALSE)
  )
  cat("\n\n")

  # Lake trout plot (core-level)
  has_lkt <- any(!is.na(df_lkt$lkt_index[df_lkt$core_name == core_label]))
  if (has_lkt) {
    p_lkt <- plot_lkt(ca) +
      labs(title = paste0(core_label, " — Lake trout abundance index"))
    print(p_lkt)
    cat("\n\n")
  }

  # ── Population subsections ────────────────────────────────────────────────
  for (pop_idx in pops_in_core) {
    cat(sprintf("\n\n### %s {#%s}\n\n",
                pop_display[pop_idx], pop_anchor[pop_idx]))

    p1 <- plot_redds(pop_idx)
    p2 <- plot_flow(pop_idx)
    p3 <- plot_temp(pop_idx)
    p4 <- plot_inv(pop_idx)

    combined <- (p1 | p2) / (p3 | p4)
    print(combined)
    cat("\n\n")
  }
}
```
