---
title: "Bull Trout Multi-Population Viability Analysis"
subtitle: "TMB State-Space Ricker Model — v1.22"
date: today
author: "Bull Trout SSP Team"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    fig-width: 10
    fig-height: 6
---

<!-- =====================================================================
     COLLABORATORS: Feel free to edit any text in this file.
     Please DO NOT modify anything inside ```{r} ... ``` code blocks —
     those generate the figures and tables automatically from the model.
     The section headings and narrative paragraphs are all fair game.
     ===================================================================== -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r load-packages}
library(here)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r load-fit}
# ── Version identifiers ──────────────────────────────────────────────────────
DataVersion     <- "1_8"
ModelVersion    <- "1_22"
RunVersion      <- "1_21"
FunctionVersion <- "1_13"

SaveName <- paste0("MPVA_Fit_D", DataVersion, "_M", ModelVersion, "_R", RunVersion)

# ── Helper functions ─────────────────────────────────────────────────────────
source(here("Code", paste0("MPVA_Functions", FunctionVersion, ".R")))

# ── Load saved model fit ─────────────────────────────────────────────────────
fit <- readRDS(here("Results", paste0(SaveName, ".RDS")))
list2env(fit, envir = .GlobalEnv)

nPop  <- data_list$nPop
nCore <- data_list$nCore
pop_core_vec <- data_list$pop_core + 1L   # 1-indexed core for each pop

# ── Convenience extractors ───────────────────────────────────────────────────
get_rep <- function(pname) {
  list(
    est = fit_rep$value[names(fit_rep$value) == pname],
    se  = fit_rep$sd[names(fit_rep$value)    == pname]
  )
}
re_summ <- summary(fit_rep, select = "random")
get_re  <- function(nm) unname(re_summ[rownames(re_summ) == nm, "Estimate"])
```

---

## Overview

This document presents the model structure and key diagnostics for a hierarchical
state-space Ricker model fitted to bull trout (*Salvelinus confluentus*) redd
count time series across multiple local populations nested within core areas.
Detailed results — covariate effects, population trajectories, exposure and
sensitivity, and future viability projections — are in the
[Model Results](results.qmd) tab.

### Spatial Hierarchy

The model uses a three-level spatial structure:

| Level | Unit | N |
|-------|------|---|
| Core area | Watershed complex | `r data_list$nCore` |
| Local population | Spawning tributary | `r data_list$nPop` |
| Survey site | Redd count reach | `r data_list$nSites` |

### Data Coverage

```{r data-summary}
year_range <- range(DatYears)
n_obs      <- sum(!is.na(data_list$ReddCounts))
n_possible <- prod(dim(data_list$ReddCounts))

data.frame(
  Metric = c("Years", "Core areas", "Local populations",
             "Survey sites", "Site-year observations", "Data coverage"),
  Value  = c(
    paste(year_range[1], "–", year_range[2]),
    data_list$nCore,
    data_list$nPop,
    data_list$nSites,
    n_obs,
    paste0(round(100 * n_obs / n_possible, 1), "%")
  )
) |>
  kable(col.names = c("Metric", "Value")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Model Description

### Process Model

The latent population abundance follows a density-dependent Ricker process
with AR(1) process error:

$$\log N_{p,t} = \log N_{p,t-1} + R_{p,t} + \varepsilon_{p,t}$$

where the instantaneous growth rate is:

$$R_{p,t} = b_{0r,p} + \beta_{\text{Flow},p} \cdot \text{Flow}_{p,t-1}
           + \beta_{\text{Temp},p} \cdot T_{p,t-1}
           + \beta_{\text{Temp}_2} \cdot T_{p,t-1}^2
           + \beta_{\text{INV},p} \cdot \text{INV}_{p,t-1}
           + \beta_{\text{LKT},M} \cdot \text{LKT}_{c,t-1}
           - \phi_p \cdot D_{p,t-1}$$

Process errors follow an AR(1) structure:

$$\varepsilon_{p,t} = \rho \cdot \varepsilon_{p,t-1} + \eta_{p,t}, \quad
  \eta_{p,t} \sim \mathcal{N}\!\left(0,\; \sigma_N \sqrt{1-\rho^2}\right)$$

All dynamic covariates carry an effective lag of **6 years** (5-year R-side lag +
1-year C++ indexing), reflecting the time from in-gravel incubation to adult
redd counts.

### Observation Model

Redd counts at each survey site follow a Negative Binomial (NB2) distribution:

$$C_{s,y} \sim \text{NB2}\!\left(\mu_{s,y},\; \phi_{\text{obs}}\right), \quad
  \mu_{s,y} = 0.5 \cdot N_{p,y} \cdot \pi_s \cdot \exp(\beta_{\text{FC}} \cdot \text{Flow}_{p,y})$$

where $\phi_{\text{obs}}$ is the NB2 dispersion parameter
($\text{Var} = \mu + \mu^2/\phi_{\text{obs}}$; as $\phi_{\text{obs}} \to \infty$
the distribution collapses to Poisson), $\pi_s$ is the fraction of population
habitat surveyed at site $s$, and $\beta_{\text{FC}}$ captures the contemporaneous
same-year flow effect on redd detectability (observation model only; does not
enter the state equation).

### CPUE Sub-Model

For populations with electrofishing data, model-based CPUE (fish km⁻¹) provides
an independent abundance index:

$$\log(\text{CPUE}_i) \sim \mathcal{N}\!\left(\log N_{p,y} - \log(\text{extent}_p)
  + q_s,\; \sigma_{\text{CPUE}}\right)$$

where $q_s$ is a section-level catchability random effect that absorbs persistent
efficiency differences among electrofishing reaches.

### Hierarchical Parameter Structure

| Parameter | Hierarchy |
|-----------|-----------|
| Baseline growth rate $b_{0r}$ | Global → Core → Population |
| Density dependence $\phi$ | Global → Population |
| Flow slope $\beta_\text{Flow}$ | Global hyper-mean → Population |
| Temperature slope $\beta_\text{Temp}$ | Global hyper-mean → Population |
| Invasive slope $\beta_\text{INV}$ | Global hyper-mean → Population |
| Lake trout slope $\beta_\text{LKT}$ | Global scalar (cores with lake trout only) |
| Adfluvial access, unconfined valley bottom | Global scalar on $b_{0r}$ mean |
| W95 winter flood frequency, road density | Global scalar on $b_{0r}$ mean |

### Three-Stage Fitting

| Stage | Observation model | Fixed | Purpose |
|-------|-------------------|-------|---------|
| A | Poisson | All SDs, AR(1) ρ | Initialize latent states |
| B1 | Poisson | All SDs | Stabilise hierarchy under Poisson |
| B2 | NB2 | — (all free) | Full model — final inference |

---

## Model Selection

### Flow Lag

Bull trout have a complex life history: eggs are deposited in late summer,
incubate over winter, and juveniles spend several years rearing before returning
as adults to spawn. We therefore tested a range of lags between environmental
conditions and the observed adult redd count response. With temperature fixed at
an effective 6-year lag (the best-supported specification from earlier runs), we
searched a grid of **effective lags 4–8 years** × **moving-average windows
(none, 2-year, 3-year)** for summer streamflow (15 combinations total). All
models used the 3-stage NB2 fitting procedure; comparison is by marginal AIC.

```{r flow-lag-selection, fig.height=4.5}
fl <- readRDS(here("Results", "FlowLagSearch_2026-03-22.RDS"))

fl_plot <- fl$results_df |>
  filter(!is.na(AIC)) |>
  mutate(
    ma_label  = case_when(
      MovAvgF == 0 ~ "No moving average",
      MovAvgF == 2 ~ "2-year MA",
      MovAvgF == 3 ~ "3-year MA"
    ),
    label     = factor(label, levels = rev(label[order(AIC)])),
    fill_grp  = case_when(
      delta_AIC <= 2 ~ "Within 2 AIC (equivalent)",
      delta_AIC <= 7 ~ "Within 7 AIC (some support)",
      TRUE           ~ "> 7 AIC (little support)"
    ) |> factor(levels = c("Within 2 AIC (equivalent)",
                            "Within 7 AIC (some support)",
                            "> 7 AIC (little support)"))
  )

ggplot(fl_plot, aes(x = label, y = delta_AIC, fill = fill_grp)) +
  geom_col(alpha = 0.85, width = 0.65) +
  geom_hline(yintercept = 2, linetype = 2, color = "goldenrod",  linewidth = 0.8) +
  geom_hline(yintercept = 7, linetype = 3, color = "tomato",     linewidth = 0.8) +
  geom_text(aes(label = sprintf("Δ%.1f", delta_AIC)),
            hjust = -0.1, size = 3.2) +
  scale_fill_manual(
    values = c("Within 2 AIC (equivalent)"  = "steelblue",
               "Within 7 AIC (some support)" = "goldenrod3",
               "> 7 AIC (little support)"    = "grey60"),
    name = NULL
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.25))) +
  coord_flip() +
  labs(
    x        = NULL,
    y        = "ΔAIC relative to best model",
    title    = "Flow lag selection — ΔAIC by lag specification",
    subtitle = sprintf(
      "Temperature fixed: effective %d-yr lag, no smoothing  |  Data v%s  |  Model v%s",
      fl$TempLag + 1, fl$DataVersion, fl$ModelVersion
    ),
    caption  = "NA = model failed to converge; longer lags exceeded the time-series extent"
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```

The **6-year effective lag with no moving average** (Lag6_NoMA) has the lowest
AIC. The next-best models all involve shorter lags with multi-year smoothing,
which may be capturing a similar biological signal through averaging. Given
the clean biological interpretation — eggs deposited in year *t* incubate for
approximately one winter and juveniles rear for approximately 4–5 years before
returning as adults — an unsmoothed 6-year lag is preferred. This specification
is used in all subsequent model runs.

### Static Habitat Covariates

With the flow lag fixed, we tested all 2^4 = 16 subsets of four static habitat
scalars that shift the prior mean of the population-level baseline growth rate:

| Covariate | Abbreviation | Biological hypothesis |
|-----------|-------------|----------------------|
| Adfluvial access (binary) | Adf | Lake-connected populations have higher adult survival / energy reserves |
| Unconfined valley bottom (z-score) | VB | Wider floodplains support more redd habitat and higher carrying capacity |
| W95 winter flood frequency (z-score) | Flood | High-flow winters scour redds and reduce recruitment |
| Road density in valley bottom (z-score) | RD | Roads increase fine sediment delivery and reduce egg-to-fry survival |

Dynamic covariates (flow, temperature, invasives, lake trout) were held at the
selected lag specification throughout. Comparison is by AICc.

```{r static-covar-selection, fig.height=6}
sc <- readRDS(here("Results", "StaticCovarSelection_2026-03-23.RDS"))

sc_plot <- sc$results_df |>
  filter(!is.na(AICc)) |>
  mutate(
    label    = factor(label, levels = rev(label[order(AICc)])),
    fill_grp = case_when(
      delta_AICc_vs_best <= 2 ~ "Within 2 AICc (equivalent)",
      delta_AICc_vs_best <= 7 ~ "Within 7 AICc (some support)",
      TRUE                    ~ "> 7 AICc (little support)"
    ) |> factor(levels = c("Within 2 AICc (equivalent)",
                            "Within 7 AICc (some support)",
                            "> 7 AICc (little support)"))
  )

ggplot(sc_plot, aes(x = label, y = delta_AICc_vs_best, fill = fill_grp)) +
  geom_col(alpha = 0.85, width = 0.65) +
  geom_hline(yintercept = 2, linetype = 2, color = "goldenrod",  linewidth = 0.8) +
  geom_hline(yintercept = 7, linetype = 3, color = "tomato",     linewidth = 0.8) +
  geom_text(aes(label = sprintf("Δ%.1f  (k=%d)", delta_AICc_vs_best, k_fixed)),
            hjust = -0.05, size = 3.0) +
  scale_fill_manual(
    values = c("Within 2 AICc (equivalent)"  = "steelblue",
               "Within 7 AICc (some support)" = "goldenrod3",
               "> 7 AICc (little support)"    = "grey60"),
    name = NULL
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.30))) +
  coord_flip() +
  labs(
    x        = NULL,
    y        = "ΔAICc relative to best model",
    title    = "Static covariate model selection — all 16 subsets",
    subtitle = sprintf(
      "Flow lag %d yr, Temp lag %d yr (no smoothing)  |  Data v%s  |  n = %d site-years",
      sc$FlowLag + 1, sc$TempLag + 1, sc$DataVersion, sc$n_obs
    )
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```

```{r static-covar-table}
sc$results_df |>
  filter(!is.na(AICc)) |>
  select(label, use_Adfluvial, use_VB, use_Flood, use_RD,
         AICc, delta_AICc_vs_best, AICc_weight) |>
  mutate(
    across(c(use_Adfluvial, use_VB, use_Flood, use_RD),
           ~ ifelse(. == 1, "✓", "")),
    across(c(AICc, delta_AICc_vs_best), ~ round(., 2)),
    AICc_weight = round(AICc_weight, 3)
  ) |>
  kable(
    col.names = c("Model", "Adf", "VB", "Flood", "RD",
                  "AICc", "ΔAICc", "w"),
    align = c("l", "c", "c", "c", "c", "r", "r", "r")
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 12) |>
  row_spec(which(sc$results_df$delta_AICc_vs_best[!is.na(sc$results_df$AICc)] <= 2),
           bold = TRUE, background = "#EBF5FB")
```

**Adfluvial access** and **flood frequency** appear in every model with strong
support (ΔAICc ≤ 2); both have clear, directional effects. Valley bottom width
and road density are individually marginal but add modest support when combined
with the core two. The **full model (all four scalars)** sits at ΔAICc = 2.6 —
within conventional equivalent-support range — and is retained to preserve
biological completeness and consistency across core areas with different habitat
profiles.

---

## Model Fit

### Convergence

```{r convergence}
conv_label <- ifelse(optB$convergence == 0, "Converged (0)",
                     paste("Not converged (", optB$convergence, ")"))
max_grad   <- max(abs(fit_rep$gradient.fixed))
rho_e      <- get_rep("rho")
phi_e      <- get_rep("phi_obs")
sigN_e     <- get_rep("sigmaN")

data.frame(
  Metric = c("Convergence code", "Negative log-likelihood", "Max |gradient|",
             "AR(1) ρ", "NB2 dispersion φ_obs", "Process error σ_N"),
  Value  = c(
    conv_label,
    round(optB$objective, 2),
    formatC(max_grad, format = "e", digits = 2),
    paste0(round(rho_e$est, 3), " (SE ", round(rho_e$se, 3), ")"),
    paste0(round(phi_e$est, 3), " (SE ", round(phi_e$se, 3), ")"),
    paste0(round(sigN_e$est, 3), " (SE ", round(sigN_e$se, 3), ")")
  )
) |>
  kable(col.names = c("Metric", "Value")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

### Observed vs Fitted Redd Counts

```{r obs-fitted-scatter}
fit_plots <- plot_fit(data_list, fit_rep)
df_obsfit <- fit_plots$fit

df_obsfit |>
  ggplot(aes(x = mu, y = y_obs)) +
  geom_point(alpha = 0.3, size = 1.2, color = "grey30") +
  geom_abline(slope = 1, intercept = 0, color = "steelblue",
              linetype = 2, linewidth = 1) +
  scale_x_continuous(trans = "log1p", labels = scales::comma) +
  scale_y_continuous(trans = "log1p", labels = scales::comma) +
  labs(x = "Fitted NB2 mean (log1p scale)", y = "Observed (log1p scale)",
       title = "Observed vs Fitted Redd Counts — All Site-Years") +
  theme_bw()
```

---

## Residual Diagnostics

```{r residuals-dist}
df_resid <- fit_plots$resid

df_resid |>
  ggplot(aes(x = pearson)) +
  geom_histogram(bins = 60, fill = "steelblue", color = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = 2, color = "grey40") +
  labs(x = "Pearson Residual", y = "Count",
       title = "Residual Distribution — All Site-Years") +
  theme_bw()
```

```{r residuals-year}
df_resid |>
  mutate(year_actual = DatYears[year]) |>
  group_by(year_actual) |>
  summarise(mean_r = mean(pearson, na.rm = TRUE),
            se_r   = sd(pearson, na.rm = TRUE) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(x = year_actual, y = mean_r)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_ribbon(aes(ymin = mean_r - 1.96 * se_r,
                  ymax = mean_r + 1.96 * se_r), alpha = 0.2, fill = "steelblue") +
  geom_line(color = "steelblue", linewidth = 0.9) +
  geom_point(size = 2) +
  labs(x = "Year", y = "Mean Pearson Residual (± 95% CI)",
       title = "Mean Residual by Year — Check for Temporal Trends") +
  theme_bw()
```

---

## Process Error and Observation Dispersion

AR(1) temporal autocorrelation and NB2 overdispersion absorb variance that
would otherwise inflate covariate effects or process error. The AR(1)
captures persistent multi-year trends in population dynamics; NB2 absorbs
aggregated spawning behaviour and observer variability in redd counts.

```{r process-obs-params}
sigq_e  <- get_rep("sigma_q")
sigc_e  <- get_rep("sigma_cpue")

params_df <- data.frame(
  Parameter = c("AR(1) autocorrelation ρ",
                "Process error σ_N (innovation SD)",
                "NB2 dispersion φ_obs",
                "CPUE observation SD σ_cpue",
                "CPUE catchability SD σ_q"),
  Estimate  = c(rho_e$est, sigN_e$est, phi_e$est, sigc_e$est, sigq_e$est),
  SE        = c(rho_e$se,  sigN_e$se,  phi_e$se,  sigc_e$se,  sigq_e$se)
) |>
  mutate(
    Lower95 = Estimate - 1.96 * SE,
    Upper95 = Estimate + 1.96 * SE
  )

params_df |>
  kable(digits = 3,
        col.names = c("Parameter", "Estimate", "SE", "Lower 95%", "Upper 95%")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Why NB2? Overdispersion Diagnostic

Redd counts show strong overdispersion relative to what a Poisson observation
model would predict. The plot below compares the observed site-level variance
against mean redd count (each point = one survey site, ≥ 5 years of data) with
the theoretical variance function for Poisson ($\text{Var} = \mu$) and NB2
($\text{Var} = \mu + \mu^2/\phi_{\text{obs}}$). The data track the NB2 curve
closely across nearly three orders of magnitude in abundance; the Poisson line
consistently underpredicts variance, especially at high counts.

```{r nb2-justification, fig.height=5}
# Build site × year data frame using the already-loaded fit objects
phi_val <- phi_e$est[1]
phi_se_ <- phi_e$se[1]

logN_mat_diag <- matrix(
  fit_rep$value[names(fit_rep$value) == "logN"],
  nrow = nPop, ncol = data_list$nYears
)

df_diag <- expand.grid(site = seq_len(data_list$nSites),
                       year = seq_len(data_list$nYears)) |>
  as_tibble() |>
  mutate(
    pop   = data_list$site_pop[site] + 1L,
    y_obs = as.vector(data_list$ReddCounts),
    mu    = 0.5 * exp(logN_mat_diag[cbind(pop, year)]) * data_list$pS[site]
  ) |>
  filter(is.finite(y_obs), !is.na(y_obs), mu > 0)

# Per-site variance vs mean (at least 5 years of data per site)
site_stats <- df_diag |>
  group_by(site) |>
  summarise(obs_mean = mean(y_obs),
            obs_var  = var(y_obs),
            n        = n(),
            .groups  = "drop") |>
  filter(n >= 5, obs_mean > 0, obs_var > 0)

# Compute global overdispersion statistics for annotation
od_pois <- sum((df_diag$y_obs - df_diag$mu)^2 / df_diag$mu) / nrow(df_diag)
od_nb2  <- sum((df_diag$y_obs - df_diag$mu)^2 /
               (df_diag$mu + df_diag$mu^2 / phi_val)) / nrow(df_diag)

# Theory curves
mu_seq <- 10^seq(log10(min(site_stats$obs_mean)),
                 log10(max(site_stats$obs_mean)),
                 length.out = 300)
var_lines <- tibble(
  mu      = mu_seq,
  Poisson = mu_seq,
  NB2     = mu_seq + mu_seq^2 / phi_val
) |> pivot_longer(-mu, names_to = "model", values_to = "var")

ggplot(site_stats, aes(x = obs_mean, y = obs_var)) +
  geom_point(alpha = 0.5, size = 2, colour = "grey30") +
  geom_line(data = var_lines,
            aes(x = mu, y = var, colour = model, linetype = model),
            linewidth = 1.0) +
  scale_colour_manual(values = c(Poisson = "steelblue", NB2 = "tomato"),
                      name = NULL) +
  scale_linetype_manual(values = c(Poisson = "dashed", NB2 = "solid"),
                        name = NULL) +
  scale_x_log10(labels = scales::comma) +
  scale_y_log10(labels = scales::comma) +
  annotate("text", x = Inf, y = -Inf, hjust = 1.05, vjust = -0.5, size = 3.2,
           label = sprintf(
             "Poisson disp. = %.1f  |  NB2 disp. = %.2f  |  φ = %.2f (SE %.2f)",
             od_pois, od_nb2, phi_val, phi_se_
           ), colour = "grey30") +
  labs(
    x        = "Site mean redd count (log scale)",
    y        = "Site observed variance (log scale)",
    title    = "Observed variance vs mean — NB2 vs Poisson",
    subtitle = sprintf(
      "NB2: Var = μ + μ²/φ,  φ = %.2f (SE = %.2f);  Poisson: Var = μ",
      phi_val, phi_se_
    )
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom")
```

---

## Causal Analysis: Invasives and Bull Trout

To determine whether the negative association between non-native trout and bull
trout reflects **displacement** (invasives suppressing bull trout population
growth) or **replacement** (declining bull trout creating space for invasives),
we fit cross-lagged mixed-effects regressions using the model-estimated latent
log-abundance as the bull trout index. The displacement signal — INV(t−k)
predicting log-λ(t) — is large, negative, and highly significant across all
lags from 1 to 8 years (|t| = 6–21), with the strongest effect at a 1-year lag
and gradual decay at longer horizons as expected from a lagged life-history
response. The replacement signal — logN(t) predicting ΔINV(t+k) — is
indistinguishable from zero at every lead (|t| < 0.4, R² < 0.01%).
The asymmetry between the two directions strongly supports a causal pathway
from invasive trout to bull trout suppression, rather than the reverse.

```{r crosslag-plot, fig.height=6}
cl <- readRDS(here("Results", "CrossLag_INV_BLT_2026-03-23.RDS"))

inv2blt <- cl$results_inv2blt |>
  mutate(sig = abs(t_stat) > 2,
         panel = "INV(t\u2212k) \u2192 log\u03bb(t)   [Displacement: invasives drive bull trout decline]")

blt2inv <- cl$results_blt2inv |>
  mutate(sig = abs(t_stat) > 2,
         panel = "logN(t) \u2192 \u0394INV(t+k)   [Replacement: bull trout decline drives invasion]")

bind_rows(inv2blt, blt2inv) |>
  mutate(panel = factor(panel, levels = c(
    "INV(t\u2212k) \u2192 log\u03bb(t)   [Displacement: invasives drive bull trout decline]",
    "logN(t) \u2192 \u0394INV(t+k)   [Replacement: bull trout decline drives invasion]"
  ))) |>
  ggplot(aes(x = lag_yr, y = beta)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50", linewidth = 0.7) +
  geom_ribbon(aes(ymin = lower95, ymax = upper95, fill = sig),
              alpha = 0.18) +
  geom_line(aes(color = sig), linewidth = 0.8) +
  geom_point(aes(color = sig, fill = sig),
             shape = 21, size = 3, stroke = 0.5) +
  scale_color_manual(values = c(`FALSE` = "grey60",  `TRUE` = "steelblue"),
                     labels = c(`FALSE` = "|t| \u2264 2", `TRUE` = "|t| > 2"),
                     name = NULL) +
  scale_fill_manual(values  = c(`FALSE` = "grey60",  `TRUE` = "steelblue"),
                    labels  = c(`FALSE` = "|t| \u2264 2", `TRUE` = "|t| > 2"),
                    name    = NULL) +
  scale_x_continuous(breaks = 0:8) +
  facet_wrap(~panel, ncol = 1, scales = "free_y") +
  labs(
    x       = "Lag / lead (years)",
    y       = "Regression coefficient (± 95% CI)",
    title   = "Cross-lagged analysis: causal direction between invasives and bull trout",
    caption = paste0("Mixed-effects regression on model-estimated latent log-abundance.  ",
                     "Run date: ", cl$run_date)
  ) +
  theme_bw(base_size = 12) +
  theme(
    legend.position  = "bottom",
    strip.text       = element_text(size = 9, face = "bold"),
    panel.grid.minor = element_blank()
  )
```

---

*Generated with [Quarto](https://quarto.org) · Model code at
[github.com/timothycline/BullTroutSSP](https://github.com/timothycline/BullTroutSSP)*
