---
title: "Bull Trout Multi-Population Viability Analysis"
subtitle: "TMB State-Space Ricker Model — v1.14"
date: today
author: "Bull Trout SSP Team"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    fig-width: 10
    fig-height: 6
---

<!-- =====================================================================
     COLLABORATORS: Feel free to edit any text in this file.
     Please DO NOT modify anything inside ```{r} ... ``` code blocks —
     those generate the figures and tables automatically from the model.
     The section headings (## Overview, ## Model Description, etc.) and
     the narrative paragraphs beneath them are all fair game.
     ===================================================================== -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r load-packages}
library(here)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r load-fit}
# ── Version identifiers ──────────────────────────────────────────────────────
DataVersion     <- "1_2"
ModelVersion    <- "1_14"
RunVersion      <- "1_14"
FunctionVersion <- "1_4"

SaveName <- paste0("MPVA_Fit_D", DataVersion, "_M", ModelVersion, "_R", RunVersion)

# ── Helper functions ─────────────────────────────────────────────────────────
source(here("Code", paste0("MPVA_Functions", FunctionVersion, ".R")))

# ── Load data (population names, years, covariates) ─────────────────────────
source(here("Code", paste0("PrepareData_MPVA", DataVersion, ".R")))

# ── Load saved model fit ─────────────────────────────────────────────────────
fit <- readRDS(here("Results", paste0(SaveName, ".RDS")))
list2env(fit, envir = .GlobalEnv)

pop_core_vec <- CoreArea_Num[match(seq_len(data_list$nPop), LocalPops_Num)]
```

---

## Overview

This document presents results from a hierarchical state-space Ricker model
for bull trout (*Salvelinus confluentus*) fitted to redd count time series
across multiple local populations nested within core areas.

### Spatial Hierarchy

The model uses a three-level spatial structure:

| Level | Unit | N |
|-------|------|---|
| Core area | Watershed complex | `r data_list$nCore` |
| Local population | Spawning tributary | `r data_list$nPop` |
| Survey site | Redd count reach | `r data_list$nSites` |

### Data Coverage

```{r data-summary}
year_range <- range(DatYears)
n_obs      <- sum(!is.na(data_list$ReddCounts))
n_possible <- prod(dim(data_list$ReddCounts))

data.frame(
  Metric = c("Years", "Core areas", "Local populations",
             "Survey sites", "Site-year observations", "Data coverage"),
  Value  = c(
    paste(year_range[1], "–", year_range[2]),
    data_list$nCore,
    data_list$nPop,
    data_list$nSites,
    n_obs,
    paste0(round(100 * n_obs / n_possible, 1), "%")
  )
) |>
  kable(col.names = c("Metric", "Value")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Model Description

### Process Model

The latent population abundance follows a density-dependent Ricker process:

$$\log N_{p,t} = \log N_{p,t-1} + R_{p,t} + \varepsilon_{p,t}$$

where the instantaneous growth rate is:

$$R_{p,t} = b_{0r,p} + \beta_{\text{Flow},p} \cdot \text{Flow}_{p,t-1}
           + \beta_{\text{Temp},p} \cdot T_{p,t-1}
           + \beta_{\text{Temp}_2} \cdot T_{p,t-1}^2
           + \beta_{\text{INV},p} \cdot \text{INV}_{p,t-1}
           + \beta_{\text{LKT},c} \cdot \text{LKT}_{c,t-1}
           - \phi_p \cdot D_{p,t-1}$$

All dynamic covariates carry an effective lag of **6 years** (5-year R-side lag +
1-year C++ indexing), reflecting the time from in-gravel incubation to adult
redd counts.

### Observation Model

Redd counts at each survey site follow a Poisson distribution:

$$C_{s,y} \sim \text{Poisson}\!\left(0.5 \cdot N_{p,y} \cdot \pi_s\right)$$

where $\pi_s$ is the fraction of population habitat surveyed at site $s$.
The factor of 0.5 converts spawning adults to females (one redd per female).

### Hierarchical Parameter Structure

| Parameter | Hierarchy |
|-----------|-----------|
| Baseline growth rate $b_{0r}$ | Global → Core → Population |
| Density dependence $\phi$ | Global → Core → Population |
| Flow slope $\beta_\text{Flow}$ | Global hyper-mean → Population |
| Temperature slope $\beta_\text{Temp}$ | Global hyper-mean → Population |
| Invasive slope $\beta_\text{INV}$ | Global hyper-mean → Population |
| Lake trout slope $\beta_\text{LKT}$ | Global hyper-mean → Core area |
| Adfluvial access, valley bottom | Global scalar on $b_{0r}$ mean |
| W95 winter flood frequency, road density | Global scalar on $b_{0r}$ mean |

---

## Model Fit

```{r convergence}
conv_label <- ifelse(optB$convergence == 0, "Converged (0)",
                     paste("Not converged (", optB$convergence, ")"))
max_grad   <- max(abs(fit_rep$gradient.fixed))

data.frame(
  Metric = c("Convergence code", "Negative log-likelihood", "Max |gradient|"),
  Value  = c(conv_label, round(optB$objective, 2), formatC(max_grad, format = "e", digits = 2))
) |>
  kable(col.names = c("Metric", "Value")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Latent Abundance Trajectories

Estimated latent abundances (number of adult spawners) for each local
population, smoothed by the state-space model.

```{r latent-abundance, fig.height=12}
fit_plots <- plot_fit(data_list, fit_rep)
dfN       <- fit_plots$latent
df_obsfit <- fit_plots$fit
df_resid  <- fit_plots$resid

dfN |>
  mutate(
    year_actual = DatYears[year],
    pop_name    = U_LocalPops_In[pop],
    core_name   = U_CoreAreas_In[pop_core_vec[pop]]
  ) |>
  ggplot(aes(x = year_actual, y = N, color = core_name)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~pop_name, scales = "free_y", ncol = 4) +
  scale_color_brewer(palette = "Set2", name = "Core area") +
  labs(x = "Year", y = "Latent Abundance (N)",
       title = "Estimated Bull Trout Spawner Abundance by Population") +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 7))
```

---

## Observed vs Fitted Redd Counts

```{r obs-fitted-time, fig.height=12}
df_obsfit |>
  mutate(
    year_actual = DatYears[year],
    pop_name    = U_LocalPops_In[pop]
  ) |>
  ggplot(aes(x = year_actual)) +
  geom_point(aes(y = y_obs), alpha = 0.6, size = 1.2, color = "grey40") +
  geom_line(aes(y = mu), linewidth = 0.8, color = "steelblue") +
  facet_wrap(~pop_name, scales = "free_y", ncol = 4) +
  labs(x = "Year", y = "Redd Counts",
       title = "Observed (points) vs Fitted (line) Redd Counts",
       caption = "Blue line = Poisson mean (0.5 × N × survey fraction); grey points = observed counts") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(size = 7))
```

```{r obs-fitted-scatter}
df_obsfit |>
  ggplot(aes(x = mu, y = y_obs)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_abline(slope = 1, intercept = 0, color = "steelblue", linetype = 2, linewidth = 1) +
  scale_x_continuous(trans = "log1p", labels = scales::comma) +
  scale_y_continuous(trans = "log1p", labels = scales::comma) +
  labs(x = "Fitted (Poisson mean, log1p scale)", y = "Observed (log1p scale)",
       title = "Observed vs Fitted — All Site-Years") +
  theme_bw()
```

---

## Residual Diagnostics

```{r residuals-time, fig.height=10}
df_resid |>
  mutate(
    year_actual = DatYears[year],
    pop_name    = U_LocalPops_In[pop]
  ) |>
  ggplot(aes(x = year_actual, y = pearson)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_point(alpha = 0.5, size = 1.2) +
  geom_smooth(method = "loess", se = FALSE, color = "steelblue", linewidth = 0.7) +
  facet_wrap(~pop_name, ncol = 4) +
  labs(x = "Year", y = "Pearson Residual",
       title = "Pearson Residuals Through Time") +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(size = 7))
```

```{r residuals-dist}
df_resid |>
  ggplot(aes(x = pearson)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = 2, color = "grey40") +
  labs(x = "Pearson Residual", y = "Count",
       title = "Residual Distribution — All Site-Years") +
  theme_bw()
```

---

## Covariate Effects

### Global Scalar Effects on Baseline Growth Rate

These covariates shift the prior mean of the population-level growth rate
intercept ($b_{0r,p}$).

```{r global-scalars}
extract_fe <- function(par_name) {
  v  <- fit_rep$value[names(fit_rep$value) == par_name]
  se <- fit_rep$sd[names(fit_rep$value)    == par_name]
  if (!length(v)) return(NULL)
  data.frame(Parameter = par_name, Estimate = v[1], SE = se[1],
             Lower95 = v[1] - 1.96 * se[1],
             Upper95 = v[1] + 1.96 * se[1])
}

global_pars <- c("bAdf", "bVB", "bFlood", "bRD", "bLKT_M")
global_df   <- bind_rows(lapply(global_pars, extract_fe)) |>
  mutate(
    Label = c("Adfluvial access", "Valley bottom width (VB)",
              "W95 winter flood frequency", "Road density (valley bottom)",
              "Lake trout ratio (cross-system mean)"),
    sig = ifelse(Lower95 > 0 | Upper95 < 0, "*", "")
  )

global_df |>
  ggplot(aes(x = reorder(Label, Estimate), y = Estimate)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.25, linewidth = 0.8) +
  geom_point(size = 3, color = "steelblue") +
  geom_text(aes(label = sig, y = Upper95 + 0.02), size = 5, color = "tomato") +
  coord_flip() +
  labs(x = NULL, y = "Coefficient estimate (95% CI)",
       title = "Static Covariate Effects on Baseline Growth Rate",
       caption = "* = 95% CI excludes zero") +
  theme_bw()
```

### Thermal Performance Curve

Bull trout growth rate shows a unimodal response to water temperature
(linear + negative quadratic).

```{r thermal-curve}
if (covar_config$use_Temp == 1) {
  plot_temp_curve(fit_rep, temp_mean = temp_mean, temp_sd = temp_sd)
}
```

### Population-Level Flow Slopes

Each local population has its own sensitivity to summer streamflow,
drawn from a global hyper-distribution (dashed red line = hyper-mean).

```{r flow-slopes, fig.height=7}
if (covar_config$use_Flow == 1) {
  bFlow_vals <- fit_rep$value[names(fit_rep$value) == "bFlow_pop"]
  bFlow_ses  <- fit_rep$sd[names(fit_rep$value)    == "bFlow_pop"]
  bFlow_M    <- fit_rep$value[names(fit_rep$value) == "bFlow_M"][1]

  data.frame(
    pop_name  = U_LocalPops_In,
    core_name = U_CoreAreas_In[pop_core_vec],
    estimate  = bFlow_vals,
    se        = bFlow_ses
  ) |>
    ggplot(aes(x = reorder(pop_name, estimate), y = estimate, color = core_name)) +
    geom_hline(yintercept = 0,       linetype = 2, color = "grey50") +
    geom_hline(yintercept = bFlow_M, linetype = 3, color = "tomato", linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - 1.96 * se,
                      ymax = estimate + 1.96 * se), width = 0.4, alpha = 0.7) +
    geom_point(size = 2.5) +
    annotate("text", x = 1, y = bFlow_M,
             label = paste0("Hyper-mean = ", round(bFlow_M, 3)),
             hjust = 0, vjust = -0.6, color = "tomato", size = 3.5) +
    coord_flip() +
    scale_color_brewer(palette = "Set2", name = "Core area") +
    labs(x = NULL, y = "Flow slope (bFlow_pop)",
         title = "Summer Streamflow Effect on Growth Rate by Population") +
    theme_bw() +
    theme(legend.position = "right")
}
```

### Population-Level Temperature Slopes

```{r temp-slopes, fig.height=7}
if (covar_config$use_Temp == 1) {
  bTemp_vals <- fit_rep$value[names(fit_rep$value) == "bTemp_pop"]
  bTemp_ses  <- fit_rep$sd[names(fit_rep$value)    == "bTemp_pop"]
  bTemp_M    <- fit_rep$value[names(fit_rep$value) == "bTemp_M"][1]

  data.frame(
    pop_name  = U_LocalPops_In,
    core_name = U_CoreAreas_In[pop_core_vec],
    estimate  = bTemp_vals,
    se        = bTemp_ses
  ) |>
    ggplot(aes(x = reorder(pop_name, estimate), y = estimate, color = core_name)) +
    geom_hline(yintercept = 0,       linetype = 2, color = "grey50") +
    geom_hline(yintercept = bTemp_M, linetype = 3, color = "tomato", linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - 1.96 * se,
                      ymax = estimate + 1.96 * se), width = 0.4, alpha = 0.7) +
    geom_point(size = 2.5) +
    annotate("text", x = 1, y = bTemp_M,
             label = paste0("Hyper-mean = ", round(bTemp_M, 3)),
             hjust = 0, vjust = -0.6, color = "tomato", size = 3.5) +
    coord_flip() +
    scale_color_brewer(palette = "Set2", name = "Core area") +
    labs(x = NULL, y = "Temperature linear slope (bTemp_pop)",
         title = "August Temperature Linear Effect on Growth Rate by Population",
         caption = "Note: global quadratic (bTemp2) is negative — combined effect is unimodal") +
    theme_bw() +
    theme(legend.position = "right")
}
```

### Population-Level Invasive Species Slopes

```{r inv-slopes, fig.height=7}
if (covar_config$use_INV == 1) {
  bINV_vals <- fit_rep$value[names(fit_rep$value) == "bINV_pop"]
  bINV_ses  <- fit_rep$sd[names(fit_rep$value)    == "bINV_pop"]
  bINV_M    <- fit_rep$value[names(fit_rep$value) == "bINV_M"][1]

  data.frame(
    pop_name  = U_LocalPops_In,
    core_name = U_CoreAreas_In[pop_core_vec],
    estimate  = bINV_vals,
    se        = bINV_ses
  ) |>
    ggplot(aes(x = reorder(pop_name, estimate), y = estimate, color = core_name)) +
    geom_hline(yintercept = 0,      linetype = 2, color = "grey50") +
    geom_hline(yintercept = bINV_M, linetype = 3, color = "tomato", linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - 1.96 * se,
                      ymax = estimate + 1.96 * se), width = 0.4, alpha = 0.7) +
    geom_point(size = 2.5) +
    annotate("text", x = 1, y = bINV_M,
             label = paste0("Hyper-mean = ", round(bINV_M, 3)),
             hjust = 0, vjust = -0.6, color = "tomato", size = 3.5) +
    coord_flip() +
    scale_color_brewer(palette = "Set2", name = "Core area") +
    labs(x = NULL, y = "Invasive ratio slope (bINV_pop)",
         title = "Non-Native Trout Ratio Effect on Growth Rate by Population") +
    theme_bw() +
    theme(legend.position = "right")
}
```

### Core-Level Lake Trout Slopes

```{r lkt-slopes}
if (covar_config$use_LKT == 1) {
  bLKT_vals <- fit_rep$value[names(fit_rep$value) == "bLKT_core"]
  bLKT_ses  <- fit_rep$sd[names(fit_rep$value)    == "bLKT_core"]
  bLKT_M    <- fit_rep$value[names(fit_rep$value) == "bLKT_M"][1]

  data.frame(
    core_name = U_CoreAreas_In,
    estimate  = bLKT_vals,
    se        = bLKT_ses
  ) |>
    ggplot(aes(x = reorder(core_name, estimate), y = estimate)) +
    geom_hline(yintercept = 0,      linetype = 2, color = "grey50") +
    geom_hline(yintercept = bLKT_M, linetype = 3, color = "tomato", linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - 1.96 * se,
                      ymax = estimate + 1.96 * se), width = 0.25, linewidth = 0.8) +
    geom_point(size = 3, color = "steelblue") +
    annotate("text", x = 1, y = bLKT_M,
             label = paste0("Hyper-mean = ", round(bLKT_M, 3)),
             hjust = 0, vjust = -0.6, color = "tomato", size = 3.5) +
    coord_flip() +
    labs(x = NULL, y = "Lake trout slope (bLKT_core)",
         title = "Lake Trout Ratio Effect on Growth Rate by Core Area") +
    theme_bw()
}
```

---

## Growth Rate and Density Dependence

### Baseline Growth Rates by Population

```{r b0r-by-core, fig.height=7}
b0r_df <- data.frame(
  pop_name  = U_LocalPops_In,
  core_name = U_CoreAreas_In[pop_core_vec],
  b0r       = fit_rep$value[names(fit_rep$value) == "b0r_pop"],
  b0r_se    = fit_rep$sd[names(fit_rep$value)    == "b0r_pop"]
)

ggplot(b0r_df, aes(x = reorder(pop_name, b0r), y = b0r, color = core_name)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_errorbar(aes(ymin = b0r - 1.96 * b0r_se,
                    ymax = b0r + 1.96 * b0r_se), width = 0.4, alpha = 0.7) +
  geom_point(size = 2.5) +
  coord_flip() +
  scale_color_brewer(palette = "Set2", name = "Core area") +
  labs(x = NULL, y = "Baseline growth rate (b0r_pop)",
       title = "Population-Level Baseline Growth Rates",
       caption = "Positive = growing at mean covariate conditions; negative = declining") +
  theme_bw()
```

### Variance Components

```{r variance-components}
sigma_pars <- c(
  "sigmaN"          = "Process error (σ_N)",
  "sigma_b0r_core"  = "Growth rate: core SD",
  "sigma_b0r_pop"   = "Growth rate: pop SD",
  "sigma_b0phi_pop" = "Density dependence: pop SD",
  "sigma_bFlow"     = "Flow slope: pop SD",
  "sigma_bTemp"     = "Temp slope: pop SD",
  "sigma_bINV"      = "Invasive slope: pop SD",
  "sigma_bLKT"      = "Lake trout slope: core SD"
)

sigma_df <- lapply(names(sigma_pars), function(nm) {
  v  <- fit_rep$value[names(fit_rep$value) == nm]
  se <- fit_rep$sd[names(fit_rep$value)    == nm]
  if (!length(v)) return(NULL)
  data.frame(Label = sigma_pars[nm], Estimate = v[1], SE = se[1])
}) |> bind_rows()

sigma_df |>
  mutate(Lower95 = pmax(0, Estimate - 1.96 * SE),
         Upper95 = Estimate + 1.96 * SE) |>
  ggplot(aes(x = reorder(Label, Estimate), y = Estimate)) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.25, linewidth = 0.8) +
  geom_point(size = 3, color = "steelblue") +
  coord_flip() +
  labs(x = NULL, y = "Standard deviation estimate (95% CI)",
       title = "Variance Component Estimates") +
  theme_bw()
```

---

## Population Table

Summary of estimated baseline growth rates and density dependence for each
local population.

```{r pop-table}
phi_pop <- exp(fit_rep$value[names(fit_rep$value) == "log_b0phi_pop"]) / 1000

data.frame(
  Population  = U_LocalPops_In,
  `Core Area` = U_CoreAreas_In[pop_core_vec],
  `b0r`       = round(b0r_df$b0r, 3),
  `b0r SE`    = round(b0r_df$b0r_se, 3),
  `phi (fish/km)` = round(phi_pop, 4)
) |>
  kable(col.names = c("Population", "Core Area", "b0r", "SE", "φ (fish/km)"),
        align     = c("l", "l", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 13) |>
  column_spec(1, bold = TRUE)
```

---

*Generated with [Quarto](https://quarto.org) · Model code at
[github.com/timothycline/BullTroutSSP](https://github.com/timothycline/BullTroutSSP)*
