---
title: "Bull Trout Multi-Population Viability Analysis"
subtitle: "TMB State-Space Ricker Model - v1.22"
date: today
author: "Bull Trout SSP Team"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    fig-width: 10
    fig-height: 6
---

<!-- =====================================================================
     COLLABORATORS: Feel free to edit any text in this file.
     Please DO NOT modify anything inside ```{r} ... ``` code blocks —
     those generate the figures and tables automatically from the model.
     The section headings (## Overview, ## Model Description, etc.) and
     the narrative paragraphs beneath them are all fair game.
     ===================================================================== -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r load-packages}
library(here)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r load-fit}
# ── Version identifiers ──────────────────────────────────────────────────────
DataVersion     <- "1_8"
ModelVersion    <- "1_22"
RunVersion      <- "1_21"
FunctionVersion <- "1_13"

SaveName <- paste0("MPVA_Fit_D", DataVersion, "_M", ModelVersion, "_R", RunVersion)

# ── Helper functions ─────────────────────────────────────────────────────────
source(here("Code", paste0("MPVA_Functions", FunctionVersion, ".R")))

# ── Load saved model fit ─────────────────────────────────────────────────────
fit <- readRDS(here("Results", paste0(SaveName, ".RDS")))
list2env(fit, envir = .GlobalEnv)

nPop  <- data_list$nPop
nCore <- data_list$nCore
pop_core_vec <- data_list$pop_core + 1L   # 1-indexed core for each pop

# ── Convenience extractors ───────────────────────────────────────────────────
get_rep <- function(pname) {
  list(
    est = fit_rep$value[names(fit_rep$value) == pname],
    se  = fit_rep$sd[names(fit_rep$value)    == pname]
  )
}
re_summ <- summary(fit_rep, select = "random")
get_re  <- function(nm) unname(re_summ[rownames(re_summ) == nm, "Estimate"])
```

```{r load-qext}
qext <- readRDS(here("Results", paste0(SaveName, "_Qext.RDS")))

nYears_proj <- qext$nYears_proj
proj_years  <- qext$proj_years
N_proj_arr  <- exp(qext$logN_proj)   # nsim x nPop x nYears_proj
nSim        <- dim(N_proj_arr)[1]

# Aggregate projected N to core level: sum populations within each core,
# then compute quantiles across simulations — gives proper uncertainty
# for the summed (core-level) abundance.
df_proj_core <- bind_rows(lapply(seq_len(nCore), function(cc) {
  pops   <- which(pop_core_vec == cc)
  Nc_sim <- if (length(pops) == 1)
               matrix(N_proj_arr[, pops, ], nrow = nSim)
             else
               apply(N_proj_arr[, pops, ], c(1, 3), sum)
  Nq <- apply(Nc_sim, 2, quantile,
               probs = c(0.05, 0.25, 0.5, 0.75, 0.95), na.rm = TRUE)
  data.frame(core_name = U_CoreAreas_In[cc], year = proj_years,
             N_p05 = Nq[1,], N_p25 = Nq[2,], N_p50 = Nq[3,],
             N_p75 = Nq[4,], N_p95 = Nq[5,])
}))

# Per-population historical bands (lognormal from sdreport marginal SEs)
logN_map <- matrix(fit_rep$value[names(fit_rep$value) == "logN"],
                   nrow = nPop, ncol = data_list$nYears)
logN_sd  <- matrix(fit_rep$sd[  names(fit_rep$value) == "logN"],
                   nrow = nPop, ncol = data_list$nYears)
z90 <- qnorm(0.95); z50 <- qnorm(0.75)

df_hist_pop <- expand.grid(pop = seq_len(nPop),
                           y_idx = seq_len(data_list$nYears)) |>
  mutate(
    pop_name  = U_LocalPops_In[pop],
    core_name = U_CoreAreas_In[pop_core_vec[pop]],
    year      = DatYears[y_idx],
    N_p50     = exp(logN_map[cbind(pop, y_idx)]),
    N_p05     = exp(logN_map[cbind(pop, y_idx)] - z90 * logN_sd[cbind(pop, y_idx)]),
    N_p25     = exp(logN_map[cbind(pop, y_idx)] - z50 * logN_sd[cbind(pop, y_idx)]),
    N_p75     = exp(logN_map[cbind(pop, y_idx)] + z50 * logN_sd[cbind(pop, y_idx)]),
    N_p95     = exp(logN_map[cbind(pop, y_idx)] + z90 * logN_sd[cbind(pop, y_idx)])
  )

# Per-population projection bands (from simulation quantiles)
r1 <- qext$qext_results[["thr20"]]
df_proj_pop <- expand.grid(pop = seq_len(nPop),
                           h   = seq_len(nYears_proj)) |>
  mutate(
    pop_name  = U_LocalPops_In[pop],
    core_name = U_CoreAreas_In[pop_core_vec[pop]],
    year      = proj_years[h],
    N_p05     = r1$N_p05[cbind(pop, h)],
    N_p25     = r1$N_p25[cbind(pop, h)],
    N_p50     = r1$N_p50[cbind(pop, h)],
    N_p75     = r1$N_p75[cbind(pop, h)],
    N_p95     = r1$N_p95[cbind(pop, h)]
  )
```

---

## Overview

This document presents results from a hierarchical state-space Ricker model
for bull trout (*Salvelinus confluentus*) fitted to redd count time series
across multiple local populations nested within core areas.

### Spatial Hierarchy

The model uses a three-level spatial structure:

| Level | Unit | N |
|-------|------|---|
| Core area | Watershed complex | `r data_list$nCore` |
| Local population | Spawning tributary | `r data_list$nPop` |
| Survey site | Redd count reach | `r data_list$nSites` |

### Data Coverage

```{r data-summary}
year_range <- range(DatYears)
n_obs      <- sum(!is.na(data_list$ReddCounts))
n_possible <- prod(dim(data_list$ReddCounts))

data.frame(
  Metric = c("Years", "Core areas", "Local populations",
             "Survey sites", "Site-year observations", "Data coverage",
             "CPUE observations"),
  Value  = c(
    paste(year_range[1], "–", year_range[2]),
    data_list$nCore,
    data_list$nPop,
    data_list$nSites,
    n_obs,
    paste0(round(100 * n_obs / n_possible, 1), "%"),
    data_list$nCPUEobs
  )
) |>
  kable(col.names = c("Metric", "Value")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Model Description

### Process Model

The latent population abundance follows a density-dependent Ricker process
with AR(1) process error:

$$\log N_{p,t} = \log N_{p,t-1} + R_{p,t} + \varepsilon_{p,t}$$

where the instantaneous growth rate is:

$$R_{p,t} = b_{0r,p} + \beta_{\text{Flow},p} \cdot \text{Flow}_{p,t-1}
           + \beta_{\text{Temp},p} \cdot T_{p,t-1}
           + \beta_{\text{Temp}_2} \cdot T_{p,t-1}^2
           + \beta_{\text{INV},p} \cdot \text{INV}_{p,t-1}
           + \beta_{\text{LKT},c} \cdot \text{LKT}_{c,t-1}
           - \phi_p \cdot D_{p,t-1}$$

Process errors follow an AR(1) structure:

$$\varepsilon_{p,t} = \rho \cdot \varepsilon_{p,t-1} + \eta_{p,t}, \quad
  \eta_{p,t} \sim \mathcal{N}\!\left(0,\; \sigma_N \sqrt{1-\rho^2}\right)$$

All dynamic covariates carry an effective lag of **6 years** (5-year R-side lag +
1-year C++ indexing), reflecting the time from in-gravel incubation to adult
redd counts.

### Observation Model

Redd counts at each survey site follow a Negative Binomial (NB2) distribution:

$$C_{s,y} \sim \text{NB2}\!\left(\mu_{s,y},\; \phi_{\text{obs}}\right), \quad
  \mu_{s,y} = 0.5 \cdot N_{p,y} \cdot \pi_s$$

where $\phi_{\text{obs}}$ is the NB2 dispersion parameter
($\text{Var} = \mu + \mu^2/\phi_{\text{obs}}$; as $\phi_{\text{obs}} \to \infty$
the distribution collapses to Poisson), and $\pi_s$ is the fraction of
population habitat surveyed at site $s$.

### CPUE Sub-Model

For populations with electrofishing data, model-based CPUE (fish km⁻¹) provides
an independent abundance index:

$$\log(\text{CPUE}_i) \sim \mathcal{N}\!\left(\log N_{p,y} - \log(\text{extent}_p)
  + q_s,\; \sigma_{\text{CPUE}}\right)$$

where $q_s$ is a section-level catchability random effect that absorbs persistent
efficiency differences among electrofishing reaches.

### Hierarchical Parameter Structure

| Parameter | Hierarchy |
|-----------|-----------|
| Baseline growth rate $b_{0r}$ | Global → Core → Population |
| Density dependence $\phi$ | Global → Core → Population |
| Flow slope $\beta_\text{Flow}$ | Global hyper-mean → Population |
| Temperature slope $\beta_\text{Temp}$ | Global hyper-mean → Population |
| Invasive slope $\beta_\text{INV}$ | Global hyper-mean → Population |
| Lake trout slope $\beta_\text{LKT}$ | Global hyper-mean → Core area |
| Adfluvial access, unconfined valley bottom | Global scalar on $b_{0r}$ mean |
| W95 winter flood frequency, road density | Global scalar on $b_{0r}$ mean |

---

## Model Fit

```{r convergence}
conv_label <- ifelse(optB$convergence == 0, "Converged (0)",
                     paste("Not converged (", optB$convergence, ")"))
max_grad   <- max(abs(fit_rep$gradient.fixed))
rho_e      <- get_rep("rho")
phi_e      <- get_rep("phi_obs")

data.frame(
  Metric = c("Convergence code", "Negative log-likelihood", "Max |gradient|",
             "AR(1) ρ", "NB2 dispersion φ"),
  Value  = c(
    conv_label,
    round(optB$objective, 2),
    formatC(max_grad, format = "e", digits = 2),
    paste0(round(rho_e$est, 3), " (SE ", round(rho_e$se, 3), ")"),
    paste0(round(phi_e$est, 3), " (SE ", round(phi_e$se, 3), ")")
  )
) |>
  kable(col.names = c("Metric", "Value")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Latent Abundance Trajectories

Estimated latent spawner abundance summed to the core-area level. Individual
population trajectories are available in the Data tab.

```{r core-trajectories, fig.height=14}
fit_plots <- plot_fit(data_list, fit_rep)
dfN       <- fit_plots$latent
df_obsfit <- fit_plots$fit
df_resid  <- fit_plots$resid

# Sum MAP latent N across populations within each core area
dfN_core <- dfN |>
  mutate(
    year_actual = DatYears[year],
    core_name   = U_CoreAreas_In[pop_core_vec[pop]]
  ) |>
  group_by(core_name, year_actual) |>
  summarise(N_total = sum(N, na.rm = TRUE), .groups = "drop")

ggplot() +
  # Historical: MAP line + shaded area
  geom_area(data = dfN_core,
            aes(x = year_actual, y = N_total),
            fill = "steelblue", alpha = 0.15) +
  geom_line(data = dfN_core,
            aes(x = year_actual, y = N_total),
            color = "steelblue", linewidth = 0.9) +
  # Divider between historical and projected
  geom_vline(xintercept = DatYears[data_list$nYears] + 0.5,
             linetype = 2, color = "grey50", linewidth = 0.4) +
  # Projected: 90% CI, 50% CI, median
  geom_ribbon(data = df_proj_core,
              aes(x = year, ymin = N_p05, ymax = N_p95),
              fill = "tomato", alpha = 0.15) +
  geom_ribbon(data = df_proj_core,
              aes(x = year, ymin = N_p25, ymax = N_p75),
              fill = "tomato", alpha = 0.30) +
  geom_line(data = df_proj_core,
            aes(x = year, y = N_p50),
            color = "tomato", linewidth = 0.9) +
  facet_wrap(~core_name, scales = "free_y", ncol = 4) +
  labs(x = "Year", y = "Estimated spawners (sum of local populations)",
       title = "Estimated Bull Trout Spawner Abundance by Core Area",
       subtitle = paste0("Historical: MAP estimate (blue) | Projected ",
                         nYears_proj, " years: median + 50/90% CI (red)",
                         " | Future covariates: ", qext$future_cov_assumption)) +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(size = 8),
        panel.grid.minor = element_blank())
```

---

## Observed vs Fitted Redd Counts

```{r obs-fitted-scatter}
df_obsfit |>
  ggplot(aes(x = mu, y = y_obs)) +
  geom_point(alpha = 0.3, size = 1.2, color = "grey30") +
  geom_abline(slope = 1, intercept = 0, color = "steelblue", linetype = 2, linewidth = 1) +
  scale_x_continuous(trans = "log1p", labels = scales::comma) +
  scale_y_continuous(trans = "log1p", labels = scales::comma) +
  labs(x = "Fitted NB2 mean (log1p scale)", y = "Observed (log1p scale)",
       title = "Observed vs Fitted Redd Counts - All Site-Years") +
  theme_bw()
```

```{r obs-fitted-pop, fig.height=20}
# Population-level summaries: total observed vs fitted by year
df_obsfit |>
  mutate(
    year_actual = DatYears[year],
    pop_name    = U_LocalPops_In[pop],
    core_name   = U_CoreAreas_In[pop_core_vec[pop]]
  ) |>
  group_by(pop_name, core_name, year_actual) |>
  summarise(y_obs = sum(y_obs, na.rm = TRUE),
            mu    = sum(mu,    na.rm = TRUE),
            .groups = "drop") |>
  ggplot(aes(x = year_actual)) +
  geom_col(aes(y = y_obs), fill = "grey70", alpha = 0.6, width = 0.8) +
  geom_line(aes(y = mu), color = "steelblue", linewidth = 0.7) +
  facet_wrap(~pop_name, scales = "free_y", ncol = 5) +
  labs(x = NULL, y = "Redd count",
       title = "Observed (bars) vs Fitted (line) Redd Counts by Population") +
  theme_bw(base_size = 9) +
  theme(strip.text = element_text(size = 6),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
```

---

## Population Viability Projections

Historical estimated abundance (blue, 50% and 90% CI from model smoothing uncertainty)
extended with a `r nYears_proj`-year stochastic forward projection (red, 50% and 90% CI
from `r qext$nsim` simulations sampling parameter and process uncertainty).
Dotted line marks N = 20 adults. Future covariates held at historical mean (z-score = 0).

```{r pop-trajectories, fig.height=22, fig.width=11}
ggplot() +
  geom_ribbon(data = df_hist_pop,
              aes(x = year, ymin = N_p05, ymax = N_p95),
              fill = "steelblue", alpha = 0.15) +
  geom_ribbon(data = df_hist_pop,
              aes(x = year, ymin = N_p25, ymax = N_p75),
              fill = "steelblue", alpha = 0.30) +
  geom_line(data = df_hist_pop,
            aes(x = year, y = N_p50),
            color = "steelblue", linewidth = 0.5) +
  geom_vline(xintercept = DatYears[data_list$nYears] + 0.5,
             linetype = 2, color = "grey50", linewidth = 0.3) +
  geom_ribbon(data = df_proj_pop,
              aes(x = year, ymin = N_p05, ymax = N_p95),
              fill = "tomato", alpha = 0.15) +
  geom_ribbon(data = df_proj_pop,
              aes(x = year, ymin = N_p25, ymax = N_p75),
              fill = "tomato", alpha = 0.30) +
  geom_line(data = df_proj_pop,
            aes(x = year, y = N_p50),
            color = "tomato", linewidth = 0.6) +
  geom_hline(yintercept = 20, linetype = 3,
             color = "firebrick", linewidth = 0.4) +
  facet_wrap(~pop_name, scales = "free_y", ncol = 5) +
  labs(x = "Year", y = "Adult Abundance (N)",
       title = "Local Population Abundance - Historical and Projected") +
  theme_bw(base_size = 7) +
  theme(strip.text      = element_text(size = 5),
        axis.text.x     = element_text(size = 5, angle = 45, hjust = 1),
        panel.grid.minor = element_blank())
```

---

## Residual Diagnostics

```{r residuals-dist}
df_resid |>
  ggplot(aes(x = pearson)) +
  geom_histogram(bins = 60, fill = "steelblue", color = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = 2, color = "grey40") +
  labs(x = "Pearson Residual", y = "Count",
       title = "Residual Distribution - All Site-Years") +
  theme_bw()
```

```{r residuals-year}
# Mean residual by year — check for systematic temporal trends
df_resid |>
  mutate(year_actual = DatYears[year]) |>
  group_by(year_actual) |>
  summarise(mean_r = mean(pearson, na.rm = TRUE),
            se_r   = sd(pearson, na.rm = TRUE) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(x = year_actual, y = mean_r)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_ribbon(aes(ymin = mean_r - 1.96 * se_r,
                  ymax = mean_r + 1.96 * se_r), alpha = 0.2, fill = "steelblue") +
  geom_line(color = "steelblue", linewidth = 0.9) +
  geom_point(size = 2) +
  labs(x = "Year", y = "Mean Pearson Residual (± 95% CI)",
       title = "Mean Residual by Year - Check for Temporal Trends") +
  theme_bw()
```

---

## Process Error and Observation Dispersion

AR(1) temporal autocorrelation and NB2 overdispersion absorb variance that
would otherwise inflate covariate effects or process error. The AR(1)
captures persistent multi-year trends in population dynamics; NB2 absorbs
aggregated spawning behaviour and observer variability in redd counts.

```{r process-obs-params}
rho_e   <- get_rep("rho")
phi_e   <- get_rep("phi_obs")
sigN_e  <- get_rep("sigmaN")
sigq_e  <- get_rep("sigma_q")
sigc_e  <- get_rep("sigma_cpue")

params_df <- data.frame(
  Parameter = c("AR(1) autocorrelation ρ",
                "Process error σ_N (innovation SD)",
                "NB2 dispersion φ_obs",
                "CPUE observation SD σ_cpue",
                "CPUE catchability SD σ_q"),
  Estimate  = c(rho_e$est, sigN_e$est, phi_e$est, sigc_e$est, sigq_e$est),
  SE        = c(rho_e$se,  sigN_e$se,  phi_e$se,  sigc_e$se,  sigq_e$se)
) |>
  mutate(
    Lower95 = Estimate - 1.96 * SE,
    Upper95 = Estimate + 1.96 * SE
  )

params_df |>
  kable(digits = 3, col.names = c("Parameter", "Estimate", "SE", "Lower 95%", "Upper 95%")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r rho-phi-plot}
params_df |>
  filter(Parameter %in% c("AR(1) autocorrelation ρ", "NB2 dispersion φ_obs")) |>
  ggplot(aes(x = Parameter, y = Estimate)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.15, linewidth = 0.9) +
  geom_point(size = 4, color = "steelblue") +
  facet_wrap(~Parameter, scales = "free", ncol = 2) +
  labs(x = NULL, y = "Estimate (95% CI)",
       title = "Process Error AR(1) Autocorrelation and NB2 Observation Dispersion") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

---

## Covariate Effects

### Static Effects on Baseline Growth Rate

These covariates shift the prior mean of the population-level growth rate
intercept ($b_{0r,p}$). Adfluvial access (lake-associated populations) and
unconfined valley bottom (floodplain habitat quality) reflect structural
habitat attributes. W95 flood frequency and road density in the valley bottom
capture disturbance.

```{r global-scalars}
global_pars <- c("bAdf", "bUVB", "bFlood", "bRD")

global_df <- lapply(global_pars, function(p) {
  r <- get_rep(p)
  if (!length(r$est)) return(NULL)
  data.frame(Parameter = p, Estimate = r$est[1], SE = r$se[1])
}) |>
  bind_rows() |>
  mutate(
    Label   = c("Adfluvial access",
                "Unconfined valley bottom (z)",
                "W95 winter flood frequency (z)",
                "Road density in valley bottom (z)"),
    Lower95 = Estimate - 1.96 * SE,
    Upper95 = Estimate + 1.96 * SE,
    sig     = ifelse(Lower95 > 0 | Upper95 < 0, "*", "")
  )

global_df |>
  ggplot(aes(x = reorder(Label, Estimate), y = Estimate)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.2, linewidth = 0.8) +
  geom_point(size = 3.5, color = "steelblue") +
  geom_text(aes(label = sig, y = Upper95 + 0.01), size = 5, color = "tomato") +
  coord_flip() +
  labs(x = NULL, y = "Coefficient estimate (95% CI)",
       title = "Static Covariate Effects on Baseline Growth Rate",
       caption = "* = 95% CI excludes zero") +
  theme_bw()
```

### Thermal Performance Curve

Bull trout growth rate shows a unimodal response to water temperature
(linear + negative quadratic).

```{r thermal-curve}
if (covar_config$use_Temp == 1) {
  plot_temp_curve(fit_rep, temp_mean = temp_mean, temp_sd = temp_sd)
}
```

### Population-Level Flow Slopes

Each local population has its own sensitivity to summer streamflow,
drawn from a global hyper-distribution (dashed red line = hyper-mean).

```{r flow-slopes, fig.height=14}
if (covar_config$use_Flow == 1) {
  r <- get_rep("bFlow_pop")
  bFlow_M <- get_rep("bFlow_M")$est[1]

  data.frame(
    pop_name  = U_LocalPops_In,
    core_name = U_CoreAreas_In[pop_core_vec],
    estimate  = r$est,
    se        = r$se
  ) |>
    ggplot(aes(x = reorder(pop_name, estimate), y = estimate, color = core_name)) +
    geom_hline(yintercept = 0,       linetype = 2, color = "grey50") +
    geom_hline(yintercept = bFlow_M, linetype = 3, color = "tomato", linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - 1.96 * se,
                      ymax = estimate + 1.96 * se), width = 0.4, alpha = 0.6) +
    geom_point(size = 1.8) +
    annotate("text", x = 1, y = bFlow_M,
             label = paste0("Hyper-mean = ", round(bFlow_M, 3)),
             hjust = 0, vjust = -0.6, color = "tomato", size = 3.5) +
    coord_flip() +
    scale_color_brewer(palette = "Set3", name = "Core area") +
    labs(x = NULL, y = "Flow slope (bFlow_pop)",
         title = "Summer Streamflow Effect on Growth Rate by Population") +
    theme_bw(base_size = 9) +
    theme(legend.position = "right",
          axis.text.y = element_text(size = 7))
}
```

### Population-Level Temperature Slopes

```{r temp-slopes, fig.height=14}
if (covar_config$use_Temp == 1) {
  r <- get_rep("bTemp_pop")
  bTemp_M <- get_rep("bTemp_M")$est[1]

  data.frame(
    pop_name  = U_LocalPops_In,
    core_name = U_CoreAreas_In[pop_core_vec],
    estimate  = r$est,
    se        = r$se
  ) |>
    ggplot(aes(x = reorder(pop_name, estimate), y = estimate, color = core_name)) +
    geom_hline(yintercept = 0,       linetype = 2, color = "grey50") +
    geom_hline(yintercept = bTemp_M, linetype = 3, color = "tomato", linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - 1.96 * se,
                      ymax = estimate + 1.96 * se), width = 0.4, alpha = 0.6) +
    geom_point(size = 1.8) +
    annotate("text", x = 1, y = bTemp_M,
             label = paste0("Hyper-mean = ", round(bTemp_M, 3)),
             hjust = 0, vjust = -0.6, color = "tomato", size = 3.5) +
    coord_flip() +
    scale_color_brewer(palette = "Set3", name = "Core area") +
    labs(x = NULL, y = "Temperature linear slope (bTemp_pop)",
         title = "August Temperature Linear Effect on Growth Rate by Population",
         caption = "Global quadratic (bTemp2) is negative - combined effect is unimodal") +
    theme_bw(base_size = 9) +
    theme(legend.position = "right",
          axis.text.y = element_text(size = 7))
}
```

### Population-Level Invasive Species Slopes

```{r inv-slopes, fig.height=14}
if (covar_config$use_INV == 1) {
  r <- get_rep("bINV_pop")
  bINV_M <- get_rep("bINV_M")$est[1]

  data.frame(
    pop_name  = U_LocalPops_In,
    core_name = U_CoreAreas_In[pop_core_vec],
    estimate  = r$est,
    se        = r$se
  ) |>
    ggplot(aes(x = reorder(pop_name, estimate), y = estimate, color = core_name)) +
    geom_hline(yintercept = 0,      linetype = 2, color = "grey50") +
    geom_hline(yintercept = bINV_M, linetype = 3, color = "tomato", linewidth = 0.8) +
    geom_errorbar(aes(ymin = estimate - 1.96 * se,
                      ymax = estimate + 1.96 * se), width = 0.4, alpha = 0.6) +
    geom_point(size = 1.8) +
    annotate("text", x = 1, y = bINV_M,
             label = paste0("Hyper-mean = ", round(bINV_M, 3)),
             hjust = 0, vjust = -0.6, color = "tomato", size = 3.5) +
    coord_flip() +
    scale_color_brewer(palette = "Set3", name = "Core area") +
    labs(x = NULL, y = "Invasive ratio slope (bINV_pop)",
         title = "Non-Native Trout Ratio Effect on Growth Rate by Population") +
    theme_bw(base_size = 9) +
    theme(legend.position = "right",
          axis.text.y = element_text(size = 7))
}
```

### Global Lake Trout Slope

In model v1.22 the lake trout effect is a single global scalar $\beta_{\text{LKT},M}$
applied to all core areas with confirmed lake trout presence (replacing the per-core
random effects used in earlier versions).

```{r lkt-slopes}
if (covar_config$use_LKT == 1) {
  r <- get_rep("bLKT_M")
  data.frame(Parameter = "bLKT_M (global scalar)",
             Estimate  = round(r$est[1], 4),
             SE        = round(r$se[1],  4),
             Lower95   = round(r$est[1] - 1.96 * r$se[1], 4),
             Upper95   = round(r$est[1] + 1.96 * r$se[1], 4)) |>
    kable(col.names = c("Parameter", "Estimate", "SE", "Lower 95%", "Upper 95%")) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
}
```

---

## Growth Rate and Density Dependence

### Baseline Growth Rates by Population

```{r b0r-by-pop, fig.height=14}
b0r <- get_rep("b0r_pop")
b0r_df <- data.frame(
  pop_name  = U_LocalPops_In,
  core_name = U_CoreAreas_In[pop_core_vec],
  b0r       = b0r$est,
  b0r_se    = b0r$se
)

ggplot(b0r_df, aes(x = reorder(pop_name, b0r), y = b0r, color = core_name)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey50") +
  geom_errorbar(aes(ymin = b0r - 1.96 * b0r_se,
                    ymax = b0r + 1.96 * b0r_se), width = 0.4, alpha = 0.6) +
  geom_point(size = 1.8) +
  coord_flip() +
  scale_color_brewer(palette = "Set3", name = "Core area") +
  labs(x = NULL, y = "Baseline growth rate (b0r_pop)",
       title = "Population-Level Baseline Growth Rates",
       caption = "Positive = growing at mean covariate conditions; negative = declining") +
  theme_bw(base_size = 9) +
  theme(axis.text.y = element_text(size = 7))
```

### Variance Components

```{r variance-components}
sigma_pars <- c(
  "sigmaN"          = "Process error σ_N (innovation)",
  "sigma_b0r_core"  = "Growth rate: core-level SD",
  "sigma_b0r_pop"   = "Growth rate: pop-level SD",
  "sigma_b0phi_pop" = "Density dependence: pop SD",
  "sigma_bFlow"     = "Flow slope: pop SD",
  "sigma_bTemp"     = "Temp slope: pop SD",
  "sigma_bINV"      = "Invasive slope: pop SD",
  "sigma_bLKT"      = "Lake trout slope: core SD",
  "sigma_cpue"      = "CPUE observation SD",
  "sigma_q"         = "CPUE catchability SD"
)

sigma_df <- lapply(names(sigma_pars), function(nm) {
  r <- get_rep(nm)
  if (!length(r$est)) return(NULL)
  data.frame(Label = sigma_pars[nm], Estimate = r$est[1], SE = r$se[1])
}) |>
  bind_rows()

sigma_df |>
  mutate(Lower95 = pmax(0, Estimate - 1.96 * SE),
         Upper95 = Estimate + 1.96 * SE) |>
  ggplot(aes(x = reorder(Label, Estimate), y = Estimate)) +
  geom_errorbar(aes(ymin = Lower95, ymax = Upper95), width = 0.25, linewidth = 0.8) +
  geom_point(size = 3, color = "steelblue") +
  coord_flip() +
  labs(x = NULL, y = "Standard deviation estimate (95% CI)",
       title = "Variance Component Estimates") +
  theme_bw()
```

---

## Population Summary Table

Estimated baseline growth rate and density dependence for each local population.

```{r pop-table}
phi_pop <- exp(fit_rep$value[names(fit_rep$value) == "log_b0phi_pop"]) / 1000

data.frame(
  Population  = U_LocalPops_In,
  `Core Area` = U_CoreAreas_In[pop_core_vec],
  `b0r`       = round(b0r_df$b0r, 3),
  `b0r SE`    = round(b0r_df$b0r_se, 3),
  `phi (fish/km)` = round(phi_pop, 4),
  check.names = FALSE
) |>
  kable(col.names = c("Population", "Core Area", "b0r", "SE", "φ (fish/km)"),
        align     = c("l", "l", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 13) |>
  column_spec(1, bold = TRUE)
```

---

*Generated with [Quarto](https://quarto.org) · Model code at
[github.com/timothycline/BullTroutSSP](https://github.com/timothycline/BullTroutSSP)*
