---
title: "Core Area Status"
subtitle: "Viability and Portfolio Metrics Across the Columbia Headwaters Recovery Unit"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load}
library(here)
library(tidyverse)
library(reactable)

DataVersion  <- "1_8"
ModelVersion <- "1_22"
OutName      <- paste0("MetapopSummary_D", DataVersion, "_M", ModelVersion)

ms   <- readRDS(here("Results", paste0(OutName, ".RDS")))
cs   <- ms$core_summary
pt   <- ms$pop_table
meta <- ms$meta

# ── Portfolio metrics computed from pop_table ──────────────────────────────
shannon_entropy <- function(x) {
  x <- x[!is.na(x) & x > 0]
  if (length(x) == 0) return(0)
  p <- x / sum(x)
  -sum(p * log(p))
}

portfolio <- pt |>
  group_by(core) |>
  summarise(
    n_pops_total = n(),
    n_occupied   = sum(N_med >= 20, na.rm = TRUE),
    pct_occupied = round(100 * n_occupied / n_pops_total, 0),
    eff_n_pops   = exp(shannon_entropy(N_med)),
    .groups      = "drop"
  ) |>
  mutate(eff_pops_rel = round(eff_n_pops / n_pops_total, 3))

cs <- cs |>
  left_join(
    select(portfolio, core, n_occupied, pct_occupied, eff_n_pops, eff_pops_rel),
    by = "core"
  )
```

## Overview

`r meta$nCore` core areas, `r meta$nPop` local populations, `r min(meta$DatYears)`–`r max(meta$DatYears)`.
Model v`r meta$ModelVersion` / Data v`r meta$DataVersion`.

Core areas approximate probable metapopulations but boundaries also reflect watershed
structure and jurisdictional units. They are the primary viability assessment unit for this
recovery unit.

## Composite Metrics: Multiple Perspectives

No single metric captures core area status fully. The five metrics used here each
prioritise a different conservation objective and are not interchangeable.

| Metric | Conservation objective | Key limitation |
|---|---|---|
| **Abundance (N/km)** | Maximise total fish | Blind to spatial distribution |
| **Growth rate (λ)** | Identify recovering vs. declining units | Backward-looking |
| **Occupancy** | Maintain all extant local populations (≥ 20 adults) | Insensitive to population size |
| **Portfolio balance** | Detect over-concentration in a single dominant population | Independent of total abundance |
| **Persistence (1 − Qext)** | Long-term viability under current trajectory | Projection-dependent |

Occupancy and portfolio balance deserve particular attention. A core that has lost
populations has reduced network redundancy regardless of total abundance — remaining
populations carry the full burden of demographic rescue. Portfolio balance (effective
populations / total, from Shannon diversity of the abundance distribution) flags cores
where fish are concentrated in one tributary: functionally a single population even if
others nominally persist.

The heatmap below presents all five metrics simultaneously, each standardised to the
range across core areas, so managers can apply their own weighting rather than a
fixed composite.

---

## Multi-Metric Status Overview

```{r heatmap, fig.height=8, fig.width=9.5}
metric_levels <- c(
  "Occupancy (%)",
  "Portfolio Balance",
  "Abundance (N/km)",
  "Growth (λ)",
  "Persistence"
)

hm_raw <- cs |>
  mutate(core_label = gsub("([a-z])([A-Z])", "\\1 \\2", core)) |>
  transmute(
    core_label,
    `Occupancy (%)`   = pct_occupied,
    `Portfolio Balance` = eff_pops_rel,
    `Abundance (N/km)` = density_N_km,
    `Growth (λ)`       = lambda_core,
    `Persistence`      = 1 - metapop_qext   # higher = better
  ) |>
  pivot_longer(-core_label, names_to = "metric", values_to = "raw") |>
  group_by(metric) |>
  mutate(
    std = (raw - min(raw, na.rm = TRUE)) /
          pmax(max(raw, na.rm = TRUE) - min(raw, na.rm = TRUE), 1e-9)
  ) |>
  ungroup() |>
  mutate(
    metric = factor(metric, levels = metric_levels),
    # Label: show the interpretable raw value in each tile
    label = case_when(
      metric == "Occupancy (%)"    ~ paste0(round(raw), "%"),
      metric == "Portfolio Balance" ~ sprintf("%.2f", raw),
      metric == "Abundance (N/km)" ~ sprintf("%.1f", raw),
      metric == "Growth (λ)"       ~ sprintf("%.3f", raw),
      metric == "Persistence"      ~ scales::percent(1 - raw, accuracy = 1),  # show Qext
      TRUE ~ as.character(raw)
    )
  )

# Order core areas by mean standardised score — best at top
core_order <- hm_raw |>
  group_by(core_label) |>
  summarise(mean_std = mean(std, na.rm = TRUE), .groups = "drop") |>
  arrange(mean_std) |>
  pull(core_label)

hm_raw <- hm_raw |>
  mutate(core_label = factor(core_label, levels = core_order))

ggplot(hm_raw, aes(x = metric, y = core_label, fill = std)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = label), size = 2.7, color = "grey10") +
  scale_fill_distiller(
    palette   = "RdYlGn",
    direction = 1,
    limits    = c(0, 1),
    name      = "Relative\nstatus",
    labels    = scales::percent_format(accuracy = 1)
  ) +
  scale_x_discrete(position = "top") +
  labs(
    x        = NULL,
    y        = NULL,
    subtitle = paste0(
      "Each metric standardised to the range across all core areas (green = best, red = worst within dataset).\n",
      "Persistence tile labels show Qext probability (lower = greener). ",
      "Portfolio Balance = effective populations / total populations.\n",
      "Core areas ordered top-to-bottom by mean rank across all five metrics."
    )
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x     = element_text(angle = 0, hjust = 0.5, face = "bold", size = 10),
    axis.text.y     = element_text(size = 9),
    panel.grid      = element_blank(),
    legend.position = "right",
    plot.subtitle   = element_text(color = "grey40", size = 8.5)
  )
```

---

## Abundance Density by Core Area

```{r density-plot, fig.height=7, fig.width=9}
cs_plot <- cs |>
  arrange(density_N_km) |>
  mutate(
    core_label = gsub("([a-z])([A-Z])", "\\1 \\2", core),
    core_label = factor(core_label, levels = unique(core_label)),
    status     = ifelse(lambda_core >= 1, "λ ≥ 1", "λ < 1")
  )

ggplot(cs_plot, aes(x = density_N_km, y = core_label, fill = status)) +
  geom_col(alpha = 0.85) +
  geom_errorbarh(
    aes(xmin = density_lo_km, xmax = density_hi_km),
    height = 0.35, linewidth = 0.5, color = "grey30"
  ) +
  scale_fill_manual(
    values = c("λ ≥ 1" = "#1565C0", "λ < 1" = "#C62828"),
    name   = "Core growth rate"
  ) +
  labs(
    x        = "Estimated abundance per km connected habitat (N/km)",
    y        = NULL,
    subtitle = "Bars = 90% CI on current N estimate"
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position  = "bottom",
    panel.grid.minor = element_blank(),
    plot.subtitle    = element_text(color = "grey40", size = 10)
  )
```

---

## Risk Matrix: Core Growth Rate vs. Quasi-Extinction Risk

```{r risk-matrix, fig.height=6, fig.width=8}
cs_risk <- cs |>
  mutate(
    core_label  = gsub("([a-z])([A-Z])", "\\1 \\2", core),
    risk_status = case_when(
      lambda_core >= 1 & metapop_qext <= 0.2  ~ "Stable",
      lambda_core >= 1 & metapop_qext >  0.2  ~ "Growing but at-risk",
      lambda_core <  1 & metapop_qext <= 0.2  ~ "Declining, lower risk",
      TRUE                                     ~ "Declining, high risk"
    ),
    risk_status = factor(risk_status,
                         levels = c("Stable", "Growing but at-risk",
                                    "Declining, lower risk", "Declining, high risk"))
  )

ggplot(cs_risk,
       aes(x = lambda_core, y = metapop_qext,
           size = hab_km, color = risk_status,
           label = core_label)) +
  annotate("rect", xmin = -Inf, xmax = 1, ymin = 0.2, ymax = Inf,
           fill = "#FFCDD2", alpha = 0.3) +
  annotate("rect", xmin = 1, xmax = Inf, ymin = 0.2, ymax = Inf,
           fill = "#FFF9C4", alpha = 0.4) +
  annotate("rect", xmin = -Inf, xmax = 1, ymin = -Inf, ymax = 0.2,
           fill = "#FFECB3", alpha = 0.3) +
  annotate("rect", xmin = 1, xmax = Inf, ymin = -Inf, ymax = 0.2,
           fill = "#C8E6C9", alpha = 0.3) +
  geom_vline(xintercept = 1, linetype = 2, color = "grey40") +
  geom_hline(yintercept = 0.2, linetype = 2, color = "grey40") +
  geom_point(alpha = 0.85) +
  ggrepel::geom_text_repel(size = 2.8, max.overlaps = 20, show.legend = FALSE) +
  scale_color_manual(
    values = c(
      "Stable"                = "#1B5E20",
      "Growing but at-risk"   = "#F9A825",
      "Declining, lower risk" = "#E65100",
      "Declining, high risk"  = "#B71C1C"
    ),
    name = NULL
  ) +
  scale_size_continuous(name = "Habitat (km)", range = c(2, 10)) +
  scale_x_continuous(labels = function(x) sprintf("%.3f", x)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x        = "Core-area intrinsic growth rate (λ)",
    y        = "Metapopulation quasi-extinction probability",
    subtitle = "Point size = connected habitat (km)  |  Qext threshold = 20 adults per population"
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position  = "bottom",
    panel.grid.minor = element_blank(),
    plot.subtitle    = element_text(color = "grey40", size = 10)
  )
```

*Quadrants: green = stable (λ ≥ 1, low Qext); yellow = growing but small or isolated;
orange = declining but with larger population buffer; red = declining with high extinction risk.*

---

## Population Status by Core Area

```{r pop-status, fig.height=7, fig.width=9}
cs_status <- cs |>
  mutate(
    core_label  = gsub("([a-z])([A-Z])", "\\1 \\2", core),
    core_label  = fct_reorder(core_label, pct_declining)
  ) |>
  select(core_label, n_pops, n_growing, n_declining) |>
  pivot_longer(c(n_growing, n_declining),
               names_to = "status", values_to = "count") |>
  mutate(
    status = recode(status,
                    n_growing   = "Growing (λ ≥ 1)",
                    n_declining = "Declining (λ < 1)")
  )

ggplot(cs_status, aes(x = count, y = core_label, fill = status)) +
  geom_col(position = "stack", alpha = 0.85) +
  scale_fill_manual(
    values = c("Growing (λ ≥ 1)" = "#1565C0", "Declining (λ < 1)" = "#C62828"),
    name   = NULL
  ) +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    x        = "Number of local populations",
    y        = NULL,
    subtitle = "Sorted by % declining (bottom = highest concern)"
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position  = "bottom",
    panel.grid.minor = element_blank(),
    plot.subtitle    = element_text(color = "grey40", size = 10)
  )
```

---

## Core-Area Summary Table

```{r summary-table}
lambda_style <- function(value) {
  if (is.na(value)) return(list())
  if (value >= 1) {
    intensity <- min(1, (value - 1) / 0.15)
    bg <- rgb(1 - intensity * 0.6, 1 - intensity * 0.4, 1)
  } else {
    intensity <- min(1, (1 - value) / 0.15)
    bg <- rgb(1, 1 - intensity * 0.7, 1 - intensity * 0.7)
  }
  list(background = bg, fontWeight = "bold")
}

qext_style <- function(value) {
  if (is.na(value)) return(list())
  intensity <- min(1, value)
  bg <- rgb(1, 1 - intensity * 0.7, 1 - intensity * 0.7)
  list(background = bg)
}

occ_style <- function(value) {
  if (is.na(value)) return(list())
  intensity <- (100 - value) / 100
  bg <- rgb(1, 1 - intensity * 0.6, 1 - intensity * 0.6)
  list(background = bg)
}

cs_tbl <- cs |>
  mutate(core = gsub("([a-z])([A-Z])", "\\1 \\2", core)) |>
  select(core, n_pops, hab_km,
         density_N_km, density_lo_km, density_hi_km,
         lambda_core, lambda_lo90, lambda_hi90,
         n_growing, n_declining, pct_declining,
         pct_occupied, eff_pops_rel,
         metapop_qext, metapop_qext_maj)

reactable(
  cs_tbl,
  filterable      = TRUE,
  searchable      = TRUE,
  sortable        = TRUE,
  resizable       = TRUE,
  defaultPageSize = 30,
  highlight       = TRUE,
  columns = list(
    core             = colDef(name = "Core Area",        minWidth = 160, sticky = "left"),
    n_pops           = colDef(name = "Pops",             maxWidth = 60),
    hab_km           = colDef(name = "Habitat (km)",     format = colFormat(digits = 1), maxWidth = 90),
    density_N_km     = colDef(name = "N/km (med)",       format = colFormat(digits = 1),
                               style = function(v) list(fontWeight = "bold"), maxWidth = 90),
    density_lo_km    = colDef(name = "N/km (5%)",        format = colFormat(digits = 1), maxWidth = 80),
    density_hi_km    = colDef(name = "N/km (95%)",       format = colFormat(digits = 1), maxWidth = 80),
    lambda_core      = colDef(name = "λ (core)",         format = colFormat(digits = 3),
                               style = lambda_style,     maxWidth = 85),
    lambda_lo90      = colDef(name = "λ (5%)",           format = colFormat(digits = 3), maxWidth = 72),
    lambda_hi90      = colDef(name = "λ (95%)",          format = colFormat(digits = 3), maxWidth = 72),
    n_growing        = colDef(name = "Growing",          maxWidth = 72),
    n_declining      = colDef(name = "Declining",        maxWidth = 80),
    pct_declining    = colDef(name = "% Declining",      format = colFormat(suffix = "%"), maxWidth = 90),
    pct_occupied     = colDef(name = "Occupancy (%)",    format = colFormat(suffix = "%"),
                               style = occ_style,        maxWidth = 100),
    eff_pops_rel     = colDef(name = "Portfolio Balance", format = colFormat(digits = 2), maxWidth = 110),
    metapop_qext     = colDef(name = "Qext (total)",     format = colFormat(percent = TRUE, digits = 1),
                               style = qext_style,       maxWidth = 95),
    metapop_qext_maj = colDef(name = "Qext (majority)",  format = colFormat(percent = TRUE, digits = 1),
                               style = qext_style,       maxWidth = 105)
  ),
  columnGroups = list(
    colGroup(name = "Habitat & Abundance",
             columns = c("n_pops", "hab_km", "density_N_km", "density_lo_km", "density_hi_km")),
    colGroup(name = "Core Growth Rate",
             columns = c("lambda_core", "lambda_lo90", "lambda_hi90")),
    colGroup(name = "Population Status",
             columns = c("n_growing", "n_declining", "pct_declining")),
    colGroup(name = "Portfolio",
             columns = c("pct_occupied", "eff_pops_rel")),
    colGroup(name = "Quasi-Extinction",
             columns = c("metapop_qext", "metapop_qext_maj"))
  )
)
```

*Occupancy = % of local populations with estimated N ≥ 20 adults (last data year).
Portfolio Balance = effective number of populations / total populations, derived from the
Shannon diversity of the within-core abundance distribution (1 = perfectly balanced; values
near 0 indicate dominance by a single population).
Qext (total) = P(total core N < 20 × n_pops at any projection year).
Qext (majority) = P(majority of populations fall below 20 adults at any projection year).
Model v`r meta$ModelVersion` | Data v`r meta$DataVersion` |
Run date: `r meta$run_date`*
