---
title: "Model Outputs"
subtitle: "Population Trajectories and Estimates by Core Area"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: right
    toc-title: "Core Areas & Populations"
---

<!-- =====================================================================
     COLLABORATORS: Feel free to edit the introductory text below.
     Please DO NOT modify anything inside ```{r} ... ``` code blocks —
     those generate all the plots and tables automatically from the model.
     ===================================================================== -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r load, include=FALSE}
library(here)
library(tidyverse)
library(patchwork)
library(knitr)
library(kableExtra)
library(scales)

# ── Version identifiers ───────────────────────────────────────────────────────
DataVersion     <- "1_6"
ModelVersion    <- "1_22"
RunVersion      <- "1_20"
FunctionVersion <- "1_13"

SaveName <- paste0("MPVA_Fit_D", DataVersion, "_M", ModelVersion, "_R", RunVersion)

source(here("Code", paste0("MPVA_Functions", FunctionVersion, ".R")))
suppressMessages(source(here("Code", paste0("PrepareData_MPVA", DataVersion, ".R"))))

fit <- readRDS(here("Results", paste0(SaveName, ".RDS")))
list2env(fit, envir = .GlobalEnv)

pop_core_vec <- CoreArea_Num[match(seq_len(data_list$nPop), LocalPops_Num)]

# ── Population display names (same logic as data.qmd) ────────────────────────
wb_raw      <- sapply(PatchWB_In, function(x) strsplit(x, "\\.")[[1]][2])
wb_readable <- trimws(gsub("([a-z0-9'])([A-Z])", "\\1 \\2", wb_raw))
pop_display <- sapply(seq_len(nPop), function(p) {
  first_site <- which(LocalPops_Num == p)[1]
  paste0(wb_readable[first_site], " (", U_LocalPops_In[p], ")")
})
pop_anchor  <- tolower(gsub("[^a-zA-Z0-9]+", "-", U_LocalPops_In))
core_anchor <- tolower(gsub("[^a-zA-Z0-9]+", "-", U_CoreAreas_In))

# ── Pre-extract logN matrix and SE matrix (ADREPORT delta-method) ────────────
logN_vals    <- fit_rep$value[names(fit_rep$value) == "logN"]
logN_se_vals <- fit_rep$sd[names(fit_rep$value)    == "logN"]
logN_mat     <- matrix(logN_vals,    nrow = nPop, ncol = nYears)
logN_se_mat  <- matrix(logN_se_vals, nrow = nPop, ncol = nYears)

# ── Pre-extract population-level sensitivity parameters ───────────────────────
bFlow_pop_vals <- fit_rep$value[names(fit_rep$value) == "bFlow_pop"]
bFlow_pop_se   <- fit_rep$sd[names(fit_rep$value)    == "bFlow_pop"]
bTemp_pop_vals <- fit_rep$value[names(fit_rep$value) == "bTemp_pop"]
bTemp_pop_se   <- fit_rep$sd[names(fit_rep$value)    == "bTemp_pop"]
bINV_pop_vals  <- fit_rep$value[names(fit_rep$value) == "bINV_pop"]
bINV_pop_se    <- fit_rep$sd[names(fit_rep$value)    == "bINV_pop"]
bFlow_M_val    <- fit_rep$value[names(fit_rep$value) == "bFlow_M"][1]
bTemp_M_val    <- fit_rep$value[names(fit_rep$value) == "bTemp_M"][1]
bINV_M_val     <- fit_rep$value[names(fit_rep$value) == "bINV_M"][1]

# ── Covariate exposure matrices (model-scale: z-scored, lagged) ───────────────
FlowMat <- data_list$Flow   # nPop  × nYears
TempMat <- data_list$Temp   # nPop  × nYears
INVMat  <- data_list$INV    # nPop  × nYears
LKTMat  <- data_list$LKT    # nCore × nYears

# ── Pre-compute observed totals and fitted mu per population-year ─────────────
# Needed for the observation overlay panel
site_pop_1idx <- data_list$site_pop + 1L   # 0-indexed → 1-indexed

obs_fitted_df <- tibble(
  site = rep(seq_len(data_list$nSites), times = nYears),
  year_idx = rep(seq_len(nYears), each = data_list$nSites),
  pop  = site_pop_1idx[rep(seq_len(data_list$nSites), times = nYears)],
  y_obs = as.vector(data_list$ReddCounts),
  pS    = data_list$pS[rep(seq_len(data_list$nSites), times = nYears)],
  N_est = exp(logN_mat)[cbind(
    site_pop_1idx[rep(seq_len(data_list$nSites), times = nYears)],
    rep(seq_len(nYears), each = data_list$nSites)
  )]
) |>
  mutate(
    mu      = 0.5 * N_est * pS,
    year    = DatYears[year_idx]
  )

obs_pop_year <- obs_fitted_df |>
  group_by(pop, year) |>
  summarise(
    obs_total  = if (any(!is.na(y_obs))) sum(y_obs, na.rm = TRUE) else NA_real_,
    fitted_mu  = sum(mu, na.rm = TRUE),
    .groups    = "drop"
  )

# ── Theme ─────────────────────────────────────────────────────────────────────
theme_out <- theme_bw(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.title       = element_text(size = 10))

THRESHOLD_ADULTS <- 20   # quasi-extinction threshold

# CI z-scores
Z90 <- qnorm(0.95)   # 1.645 — 90% CI
Z50 <- qnorm(0.75)   # 0.674 — 50% CI

# ── Population plot function ──────────────────────────────────────────────────
#   Single panel: 90% CI ribbon (light) + 50% CI ribbon (medium) + MLE line
#   + observed redd counts scaled to adult-equivalent as points
#   + management threshold
plot_trajectory <- function(pop_idx) {

  se <- logN_se_mat[pop_idx, ]
  lN <- logN_mat[pop_idx, ]

  traj <- tibble(
    year    = DatYears,
    N       = exp(lN),
    N_lo90  = exp(lN - Z90 * se),
    N_hi90  = exp(lN + Z90 * se),
    N_lo50  = exp(lN - Z50 * se),
    N_hi50  = exp(lN + Z50 * se)
  )

  # Scale observed redd totals to adult-equivalent: N_equiv = redd_total / (0.5 * pS_total)
  # pS_total = sum of propSite fractions for this population
  sites_p  <- which(site_pop_1idx == pop_idx)
  pS_total <- sum(data_list$pS[sites_p])

  obs <- obs_pop_year |>
    filter(pop == pop_idx) |>
    mutate(N_obs_equiv = obs_total / (0.5 * pS_total))

  # y-axis upper limit: accommodate ribbons and scaled observations
  y_max <- max(
    max(traj$N_hi90,        na.rm = TRUE),
    max(obs$N_obs_equiv,    na.rm = TRUE),
    THRESHOLD_ADULTS * 1.5
  ) * 1.08

  ggplot(traj, aes(x = year)) +
    # 90% CI — outer, lightest
    geom_ribbon(aes(ymin = N_lo90, ymax = N_hi90),
                fill = "steelblue", alpha = 0.15) +
    # 50% CI — inner, more visible
    geom_ribbon(aes(ymin = N_lo50, ymax = N_hi50),
                fill = "steelblue", alpha = 0.30) +
    # MLE trajectory
    geom_line(aes(y = N), color = "steelblue4", linewidth = 1.1) +
    # Observed redd counts scaled to adult-equivalent
    geom_point(data = obs, aes(x = year, y = N_obs_equiv),
               shape = 21, fill = "white", color = "grey30",
               size = 2.2, stroke = 0.8, na.rm = TRUE) +
    # Management threshold
    geom_hline(yintercept = THRESHOLD_ADULTS,
               linetype = 2, color = "tomato3", linewidth = 0.8) +
    annotate("text",
             x     = min(DatYears) + 0.5,
             y     = THRESHOLD_ADULTS * 1.10,
             label = paste0(THRESHOLD_ADULTS, "-adult threshold"),
             color = "tomato3", size = 3.0, hjust = 0) +
    scale_y_continuous(labels = comma, limits = c(0, y_max)) +
    labs(
      x        = NULL,
      y        = "Adults (estimated / observed equivalent)",
      title    = "Estimated adult abundance",
      subtitle = "Ribbons: 50% CI (dark) and 90% CI (light)  |  Points: observed redds ÷ (0.5 × survey coverage)"
    ) +
    theme_out
}

# ── Sensitivity forest plot ────────────────────────────────────────────────────
# Shows bFlow_pop, bTemp_pop, bINV_pop ± 95% CI for all pops in a core area.
# Gold dotted line = global hyper-mean; filled points = CI excludes zero.
plot_sensitivity <- function(pops_in_core) {

  covars <- list(
    list(vals = bFlow_pop_vals, ses = bFlow_pop_se,
         mean = bFlow_M_val,   label = "Summer flow"),
    list(vals = bTemp_pop_vals, ses = bTemp_pop_se,
         mean = bTemp_M_val,   label = "August temperature"),
    list(vals = bINV_pop_vals,  ses = bINV_pop_se,
         mean = bINV_M_val,    label = "Non-native trout")
  )

  sens_df <- bind_rows(lapply(covars, function(cv) {
    tibble(
      pop_name   = pop_display[pops_in_core],
      covariate  = cv$label,
      estimate   = cv$vals[pops_in_core],
      se         = cv$ses[pops_in_core],
      hyper_mean = cv$mean
    )
  })) |>
    mutate(
      lo95      = estimate - 1.96 * se,
      hi95      = estimate + 1.96 * se,
      sig       = (lo95 > 0) | (hi95 < 0),
      pop_name  = factor(pop_name, levels = rev(pop_display[pops_in_core])),
      covariate = factor(covariate,
                         levels = c("Summer flow", "August temperature",
                                    "Non-native trout"))
    )

  ggplot(sens_df, aes(x = estimate, y = pop_name)) +
    geom_vline(xintercept = 0, linetype = 2, color = "grey50", linewidth = 0.7) +
    geom_vline(aes(xintercept = hyper_mean),
               linetype = 3, color = "goldenrod3", linewidth = 0.9) +
    geom_errorbarh(aes(xmin = lo95, xmax = hi95),
                   height = 0.3, alpha = 0.55, color = "grey40") +
    geom_point(aes(fill = sig), shape = 21, size = 2.8, color = "grey30", stroke = 0.5) +
    scale_fill_manual(values = c(`FALSE` = "white", `TRUE` = "steelblue3"),
                      guide = "none") +
    facet_wrap(~covariate, scales = "free_x", nrow = 1) +
    labs(
      x        = "Effect on log-λ per SD covariate",
      y        = NULL,
      title    = "Population sensitivity to dynamic covariates",
      subtitle = "Error bars: 95% CI  |  Gold dotted: global hyper-mean  |  Filled: CI excludes zero"
    ) +
    theme_out +
    theme(strip.text       = element_text(size = 9, face = "bold"),
          panel.spacing.x  = unit(12, "pt"))
}

# ── Exposure time series ───────────────────────────────────────────────────────
# Shows the model-scale covariate time series for each population in a core.
# Covariates are z-scored and lagged as used by the model.
plot_exposure <- function(pops_in_core, ca) {

  yr <- DatYears
  pop_labels <- setNames(pop_display[pops_in_core], as.character(pops_in_core))

  mat_to_long <- function(mat, pops, label) {
    bind_rows(lapply(pops, function(p) {
      tibble(year = yr, value = mat[p, ], pop_name = pop_display[p])
    })) |> filter(!is.na(value) & is.finite(value))
  }

  df_f <- mat_to_long(FlowMat, pops_in_core)
  df_t <- mat_to_long(TempMat, pops_in_core)
  df_i <- mat_to_long(INVMat,  pops_in_core)

  n_pops_c   <- length(pops_in_core)
  pal        <- scales::hue_pal()(n_pops_c)
  pop_names  <- pop_display[pops_in_core]
  color_scale <- scale_color_manual(values = setNames(pal, pop_names), name = NULL)

  p_f <- ggplot(df_f, aes(x = year, y = value, color = pop_name)) +
    geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
    geom_line(linewidth = 0.75, alpha = 0.85) +
    color_scale +
    guides(color = "none") +
    labs(x = NULL, y = "z-score", title = "Summer streamflow (z, lag 6yr)") +
    theme_out

  p_t <- ggplot(df_t, aes(x = year, y = value, color = pop_name)) +
    geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
    geom_line(linewidth = 0.75, alpha = 0.85) +
    color_scale +
    guides(color = "none") +
    labs(x = NULL, y = "z-score", title = "August temperature (z, lag 6yr)") +
    theme_out

  p_i <- ggplot(df_i, aes(x = year, y = value, color = pop_name)) +
    geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
    geom_line(linewidth = 0.75, alpha = 0.85) +
    color_scale +
    labs(x = NULL, y = "standardized index", title = "Non-native trout") +
    theme_out +
    theme(legend.position  = "bottom",
          legend.text      = element_text(size = 7),
          legend.key.width = unit(16, "pt"))

  # LKT: core-level — show only if core has real data
  has_lkt_ca  <- isTRUE(data_list$has_LKT[ca] == 1L)
  lkt_vals    <- if (has_lkt_ca) LKTMat[ca, ] else rep(NA_real_, nYears)
  df_l <- tibble(year = yr, value = lkt_vals) |>
    filter(!is.na(value) & is.finite(value))

  p_l <- if (nrow(df_l) > 0) {
    ggplot(df_l, aes(x = year, y = value)) +
      geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
      geom_area(fill = "purple4", alpha = 0.15) +
      geom_line(color = "purple4", linewidth = 0.9) +
      labs(x = NULL, y = "standardized index", title = "Lake trout (core-level)") +
      theme_out
  } else {
    ggplot() +
      annotate("text", x = 0.5, y = 0.5, label = "No lake trout data\nfor this core area",
               size = 4, color = "grey60") +
      labs(title = "Lake trout (core-level)") +
      theme_out
  }

  (p_f | p_t) / (p_i | p_l)
}
```

This page presents model results organized identically to the
[Data](data.qmd) tab. Use the table of contents on the right to jump
directly to any core area or local population.

---

## Global Results

```{r global-fit, results='asis'}
# ── Fit summary ───────────────────────────────────────────────────────────────
max_grad <- max(abs(fit_rep$gradient.fixed))
n_obs_   <- sum(!is.na(data_list$ReddCounts))
n_par    <- length(optB$par)

tibble(
  Metric = c("Convergence code", "Objective (NLL)", "Max |gradient|",
             "Site-year observations", "Outer parameters"),
  Value  = c(
    ifelse(optB$convergence == 0, "0 (relative convergence)", optB$convergence),
    round(optB$objective, 2),
    formatC(max_grad, format = "e", digits = 2),
    n_obs_,
    n_par
  )
) |>
  kable(col.names = c("Metric", "Value"), align = c("l","r")) |>
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE, font_size = 12)
```

### Covariate Effects

```{r covar-table}
# Helper: extract estimate + SE by parameter name, return NA if absent
get_est <- function(nm) {
  v <- fit_rep$value[names(fit_rep$value) == nm]
  s <- fit_rep$sd[names(fit_rep$value)    == nm]
  if (!length(v)) return(c(NA_real_, NA_real_))
  c(v[1], s[1])
}

covar_rows <- list(
  list(par = "bFlow_M",  label = "Summer flow (hyper-mean)",       unit = "log-λ / SD flow"),
  list(par = "bTemp_M",  label = "August temperature (hyper-mean)", unit = "log-λ / SD temp"),
  list(par = "bTemp2",   label = "Temperature² (quadratic)",        unit = "log-λ / SD temp²"),
  list(par = "bINV_M",   label = "Non-native trout (hyper-mean)",   unit = "log-λ / SD INV"),
  list(par = "bLKT_M",   label = "Lake trout (global scalar)",      unit = "log-λ / SD LKT"),
  list(par = "bAdf",     label = "Lake area (adfluvial access)",    unit = "log-λ / log(km²+1)"),
  list(par = "bUVB",     label = "Unconfined valley bottom",         unit = "log-λ / z-score"),
  list(par = "bFlood",   label = "W95 flood frequency",              unit = "log-λ / z-score"),
  list(par = "bRD",      label = "Road density",                    unit = "log-λ / z-score")
)

covar_df <- bind_rows(lapply(covar_rows, function(r) {
  es <- get_est(r$par)
  tibble(Parameter = r$label, Units = r$unit,
         Estimate = es[1], SE = es[2])
})) |>
  filter(!is.na(Estimate)) |>
  mutate(
    z     = Estimate / SE,
    p     = 2 * pnorm(-abs(z)),
    Sig   = case_when(
      abs(z) >= 2.576 ~ "***",
      abs(z) >= 1.960 ~ "**",
      abs(z) >= 1.645 ~ "*",
      TRUE            ~ ""
    ),
    across(c(Estimate, SE, z), ~ round(., 3)),
    p = round(p, 4)
  )

covar_df |>
  kable(col.names = c("Covariate", "Units", "Estimate", "SE",
                       "z", "p", ""),
        align = c("l","l","r","r","r","r","l")) |>
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE, font_size = 12) |>
  footnote(general = "*** p < 0.01  ** p < 0.05  * p < 0.10.  All dynamic covariates carry an effective 6-yr lag.")
```

### Variance Components

```{r variance-table}
var_rows <- list(
  list(par = "sigmaN",          label = "Process noise (σ_N)",              type = "Process"),
  list(par = "rho",             label = "AR(1) autocorrelation (ρ)",         type = "Process"),
  list(par = "phi_obs",         label = "NB2 dispersion (φ_obs)",            type = "Observation"),
  list(par = "sigma_b0r_pop",   label = "Growth rate SD — population",       type = "Hierarchy"),
  list(par = "sigma_b0r_core",  label = "Growth rate SD — core area",        type = "Hierarchy"),
  list(par = "sigma_b0phi_pop", label = "Density dependence SD — population",type = "Hierarchy"),
  list(par = "sigma_bFlow",     label = "Flow slope SD — population",        type = "Hierarchy"),
  list(par = "sigma_bTemp",     label = "Temp slope SD — population",        type = "Hierarchy"),
  list(par = "sigma_bINV",      label = "Invasives slope SD — population",   type = "Hierarchy")
)

var_df <- bind_rows(lapply(var_rows, function(r) {
  es <- get_est(r$par)
  tibble(Type = r$type, Parameter = r$label,
         Estimate = round(es[1], 3), SE = round(es[2], 3))
})) |>
  filter(!is.na(Estimate))

var_df |>
  kable(col.names = c("Type", "Parameter", "Estimate", "SE"),
        align = c("l","l","r","r")) |>
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE, font_size = 12) |>
  column_spec(1, bold = TRUE)
```

---

## Population Trajectories

Each plot shows **estimated adult abundance** (blue line) with **50% CI** (dark ribbon)
and **90% CI** (light ribbon) derived from the delta method on the ADREPORT-ed
log-scale state. Open circles are observed redd counts scaled to adult-equivalents
(redds ÷ 0.5 × survey coverage fraction), placing observations and model estimates
on the same axis. The dashed red line marks the **`r THRESHOLD_ADULTS`-adult
management threshold**.

```{r core-index, results='asis'}
cat("### Jump to core area\n\n")
for (ca in seq_len(nCore)) {
  n_pops <- sum(pop_core_vec == ca)
  cat(sprintf("- [%s](#%s) — %d population%s\n",
              U_CoreAreas_In[ca], core_anchor[ca],
              n_pops, if (n_pops > 1) "s" else ""))
}
cat("\n---\n\n")
```

```{r all-sections, results='asis', fig.width=10, fig.height=7}
# Pre-extract core-level growth rates
b0r_core_vals <- fit_rep$value[names(fit_rep$value) == "b0r_core"]
b0r_core_se   <- fit_rep$sd[names(fit_rep$value)    == "b0r_core"]
b0r_pop_vals  <- fit_rep$value[names(fit_rep$value) == "b0r_pop"]
b0r_pop_se    <- fit_rep$sd[names(fit_rep$value)    == "b0r_pop"]

for (ca in seq_len(nCore)) {

  core_label   <- U_CoreAreas_In[ca]
  pops_in_core <- which(pop_core_vec == ca)
  n_pops       <- length(pops_in_core)

  # Core header
  cat(sprintf("\n\n## %s {#%s}\n\n", core_label, core_anchor[ca]))

  # Core data year range
  yr_range_core <- range(DatYears[apply(
    data_list$ReddCounts[which(site_pop_1idx %in% pops_in_core), , drop = FALSE],
    2, function(x) any(!is.na(x))
  ) |> which()])

  cat(sprintf("**%d local population%s** | Data span: **%d – %d**\n\n",
              n_pops, if (n_pops > 1) "s" else "",
              yr_range_core[1], yr_range_core[2]))

  # Core-level growth rate table
  core_b0r <- b0r_core_vals[ca]
  core_b0r_se <- b0r_core_se[ca]
  if (!is.na(core_b0r)) {
    cat(sprintf(
      "> **Core growth rate** (b0r_core): **%.3f** (SE: %.3f) — lambda ≈ **%.3f**\n\n",
      core_b0r, core_b0r_se, exp(core_b0r)
    ))
  }

  # Population growth rate summary table for this core
  pop_summary <- tibble(
    Population   = pop_display[pops_in_core],
    `b0r (pop)`  = round(b0r_pop_vals[pops_in_core], 3),
    `SE`         = round(b0r_pop_se[pops_in_core],   3),
    `λ (exp b0r)` = round(exp(b0r_pop_vals[pops_in_core]), 3)
  )
  print(
    kable(pop_summary, align = c("l","r","r","r")) |>
    kable_styling(bootstrap_options = c("striped","condensed"),
                  font_size = 11, full_width = FALSE)
  )
  cat("\n\n")

  # ── Sensitivity: population-level covariate slopes ──────────────────────────
  cat("#### Covariate sensitivity\n\n")
  sens_height <- max(2.5, 0.45 * n_pops + 1.2)
  print(plot_sensitivity(pops_in_core))
  cat("\n\n")

  # ── Exposure: covariate time series for this core ────────────────────────────
  cat("#### Covariate exposure\n\n")
  print(plot_exposure(pops_in_core, ca))
  cat("\n\n<hr>\n\n")

  # Population subsections
  for (pop_idx in pops_in_core) {
    cat(sprintf("\n\n### %s {#%s}\n\n",
                pop_display[pop_idx], pop_anchor[pop_idx]))

    print(plot_trajectory(pop_idx))
    cat("\n\n")
  }
}
```
