---
title: "Conservation Prioritization"
subtitle: "Population Status and Scenario Explorer"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load}
library(here)
library(tidyverse)
library(reactable)

DataVersion  <- "1_8"
ModelVersion <- "1_22"
OutName      <- paste0("PrioritizationData_D", DataVersion, "_M", ModelVersion)

pdat <- readRDS(here("Results", paste0(OutName, ".RDS")))
st   <- pdat$status_table
meta <- pdat$meta
```

## Overview

This page presents the demographic status of all `r meta$nPop` local bull trout
populations across `r length(meta$core_names)` core areas in the Columbia
Headwaters Recovery Unit. Estimates are derived from a hierarchical state-space
Ricker model fitted to annual redd count time series
(Model v`r meta$ModelVersion` / Data v`r meta$DataVersion`).

**Interpreting the table:**

- **λ (median)** — intrinsic growth rate under current covariate conditions
  (last observed year). Values > 1 indicate a growing population; < 1 declining.
- **Qext (N < 20)** — probability of dropping below 20 adults at any point in
  a 7-year forward projection under mean environmental conditions.
- **Sensitivity columns** — λ if a single threat lever is improved while all
  others remain at current levels. These highlight which interventions offer
  the greatest potential benefit per population.

---

## Interactive Scenario Tool

The scenario tool lets you set covariate conditions across all populations
simultaneously and see how projected λ changes — useful for comparing the
relative benefit of different management actions and identifying which
populations respond most strongly to each threat lever.

```{=html}
<div style="margin: 16px 0 24px 0;">
  <a href="https://PLACEHOLDER.shinyapps.io/scenario_app/"
     class="btn btn-primary"
     target="_blank"
     style="font-size: 15px; padding: 10px 24px;">
    &#9654;&nbsp; Launch Scenario Explorer
  </a>
  <span style="margin-left: 12px; color: #666; font-size: 13px;">
    Opens in a new tab
  </span>
</div>
```

> **Note:** Replace `PLACEHOLDER` above with your shinyapps.io account name
> after deploying with `rsconnect::deployApp("ShinyApp/scenario_app")`.

---

## Population Status Table

```{r status-table}
lambda_style <- function(value) {
  if (is.na(value)) return(list())
  if (value >= 1) {
    intensity <- min(1, (value - 1) / 0.5)
    bg <- rgb(1 - intensity * 0.6, 1 - intensity * 0.4, 1)
  } else {
    intensity <- min(1, (1 - value) / 0.5)
    bg <- rgb(1, 1 - intensity * 0.7, 1 - intensity * 0.7)
  }
  list(background = bg, fontWeight = "bold")
}

qext_style <- function(value) {
  if (is.na(value)) return(list())
  intensity <- min(1, value)
  bg <- rgb(1, 1 - intensity * 0.7, 1 - intensity * 0.7)
  list(background = bg)
}

reactable(
  st |> select(core, pop_name, pop_id,
               lambda_med, lambda_lo90, lambda_hi90,
               qext_p20,
               lam_inv_mean, lam_inv_best,
               lam_temp_1sd, lam_flow_1sd, lam_lkt_best),
  filterable      = TRUE,
  searchable      = TRUE,
  sortable        = TRUE,
  resizable       = TRUE,
  defaultPageSize = 30,
  highlight       = TRUE,
  columns = list(
    core         = colDef(name = "Core Area",    minWidth = 130, sticky = "left"),
    pop_name     = colDef(name = "Population",   minWidth = 120),
    pop_id       = colDef(name = "Pop ID",       maxWidth = 90),
    lambda_med   = colDef(name = "\u03bb (median)",  format = colFormat(digits = 2),
                          style = lambda_style,  minWidth = 85),
    lambda_lo90  = colDef(name = "\u03bb (5%)",  format = colFormat(digits = 2), maxWidth = 72),
    lambda_hi90  = colDef(name = "\u03bb (95%)", format = colFormat(digits = 2), maxWidth = 72),
    qext_p20     = colDef(name = "Qext (N<20)", format = colFormat(percent = TRUE, digits = 1),
                          style = qext_style,   maxWidth = 95),
    lam_inv_mean = colDef(name = "\u03bb (INV avg)",  format = colFormat(digits = 2),
                          style = lambda_style,  maxWidth = 90),
    lam_inv_best = colDef(name = "\u03bb (INV best)", format = colFormat(digits = 2),
                          style = lambda_style,  maxWidth = 90),
    lam_temp_1sd = colDef(name = "\u03bb (Temp \u22121SD)", format = colFormat(digits = 2),
                          style = lambda_style,  maxWidth = 95),
    lam_flow_1sd = colDef(name = "\u03bb (Flow +1SD)", format = colFormat(digits = 2),
                          style = lambda_style,  maxWidth = 95),
    lam_lkt_best = colDef(name = "\u03bb (LKT best)", format = colFormat(digits = 2),
                          style = lambda_style,  na = "\u2014", maxWidth = 90)
  ),
  columnGroups = list(
    colGroup(name = "Current Status",
             columns = c("lambda_med", "lambda_lo90", "lambda_hi90")),
    colGroup(name = "Quasi-Extinction",
             columns = c("qext_p20")),
    colGroup(name = "Sensitivity — best-case \u03bb per lever",
             columns = c("lam_inv_mean", "lam_inv_best",
                         "lam_temp_1sd", "lam_flow_1sd", "lam_lkt_best"))
  )
)
```

*Model v`r meta$ModelVersion` | Data v`r meta$DataVersion` |
Run date: `r meta$run_date` |
`r meta$nsamp` marginal posterior draws*
