---
title: "Model Results"
subtitle: "Population Trajectories, Covariate Effects, and Viability Projections"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: right
    toc-title: "Core Areas & Populations"
---

<!-- =====================================================================
     COLLABORATORS: Feel free to edit the introductory text below.
     Please DO NOT modify anything inside ```{r} ... ``` code blocks —
     those generate all the plots and tables automatically from the model.
     ===================================================================== -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r load, include=FALSE}
library(here)
library(tidyverse)
library(patchwork)
library(knitr)
library(kableExtra)
library(scales)

# ── Version identifiers ───────────────────────────────────────────────────────
DataVersion     <- "1_10"
ModelVersion    <- "1_22"
RunVersion      <- "1_22"
FunctionVersion <- "1_13"

SaveName <- paste0("MPVA_Fit_D", DataVersion, "_M", ModelVersion, "_R", RunVersion)

source(here("Code", paste0("MPVA_Functions", FunctionVersion, ".R")))
suppressMessages(source(here("Code", paste0("PrepareData_MPVA", DataVersion, ".R"))))

fit <- readRDS(here("Results", paste0(SaveName, ".RDS")))
list2env(fit, envir = .GlobalEnv)

pop_core_vec <- data_list$pop_core + 1L   # 1-indexed core for each pop

# ── Population display names ──────────────────────────────────────────────────
wb_raw      <- sapply(PatchWB_In, function(x) strsplit(x, "\\.")[[1]][2])
wb_readable <- trimws(gsub("([a-z0-9'])([A-Z])", "\\1 \\2", wb_raw))
pop_display <- sapply(seq_len(nPop), function(p) {
  first_site <- which(LocalPops_Num == p)[1]
  paste0(wb_readable[first_site], " (", U_LocalPops_In[p], ")")
})
pop_anchor  <- tolower(gsub("[^a-zA-Z0-9]+", "-", U_LocalPops_In))
core_anchor <- tolower(gsub("[^a-zA-Z0-9]+", "-", U_CoreAreas_In))

# ── Pre-extract logN matrix and SE matrix (ADREPORT delta-method) ────────────
logN_vals    <- fit_rep$value[names(fit_rep$value) == "logN"]
logN_se_vals <- fit_rep$sd[names(fit_rep$value)    == "logN"]
logN_mat     <- matrix(logN_vals,    nrow = nPop, ncol = nYears)
logN_se_mat  <- matrix(logN_se_vals, nrow = nPop, ncol = nYears)

# ── Pre-extract population-level sensitivity parameters ───────────────────────
bFlow_pop_vals <- fit_rep$value[names(fit_rep$value) == "bFlow_pop"]
bFlow_pop_se   <- fit_rep$sd[names(fit_rep$value)    == "bFlow_pop"]
bTemp_pop_vals <- fit_rep$value[names(fit_rep$value) == "bTemp_pop"]
bTemp_pop_se   <- fit_rep$sd[names(fit_rep$value)    == "bTemp_pop"]
bINV_pop_vals  <- fit_rep$value[names(fit_rep$value) == "bINV_pop"]
bINV_pop_se    <- fit_rep$sd[names(fit_rep$value)    == "bINV_pop"]
bFlow_M_val    <- fit_rep$value[names(fit_rep$value) == "bFlow_M"][1]
bTemp_M_val    <- fit_rep$value[names(fit_rep$value) == "bTemp_M"][1]
bINV_M_val     <- fit_rep$value[names(fit_rep$value) == "bINV_M"][1]

# ── Covariate exposure matrices (model-scale: z-scored, lagged) ───────────────
FlowMat <- data_list$Flow   # nPop  × nYears
TempMat <- data_list$Temp   # nPop  × nYears
INVMat  <- data_list$INV    # nPop  × nYears
LKTMat  <- data_list$LKT    # nCore × nYears

# ── Pre-compute observed totals and fitted mu per population-year ─────────────
site_pop_1idx <- data_list$site_pop + 1L   # 0-indexed → 1-indexed

obs_fitted_df <- tibble(
  site     = rep(seq_len(data_list$nSites), times = nYears),
  year_idx = rep(seq_len(nYears), each = data_list$nSites),
  pop      = site_pop_1idx[rep(seq_len(data_list$nSites), times = nYears)],
  y_obs    = as.vector(data_list$ReddCounts),
  pS       = data_list$pS[rep(seq_len(data_list$nSites), times = nYears)],
  N_est    = exp(logN_mat)[cbind(
    site_pop_1idx[rep(seq_len(data_list$nSites), times = nYears)],
    rep(seq_len(nYears), each = data_list$nSites)
  )]
) |>
  mutate(mu = 0.5 * N_est * pS, year = DatYears[year_idx])

obs_pop_year <- obs_fitted_df |>
  group_by(pop, year) |>
  summarise(
    obs_total = if (any(!is.na(y_obs))) sum(y_obs, na.rm = TRUE) else NA_real_,
    fitted_mu = sum(mu, na.rm = TRUE),
    .groups   = "drop"
  )

# ── Theme ─────────────────────────────────────────────────────────────────────
theme_out <- theme_bw(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.title       = element_text(size = 10))

THRESHOLD_ADULTS <- 20   # demographic risk threshold
Z90 <- qnorm(0.95)       # 1.645 — 90% CI
Z50 <- qnorm(0.75)       # 0.674 — 50% CI
```

```{r load-qext, include=FALSE}
qext <- readRDS(here("Results", paste0(SaveName, "_Qext.RDS")))

nYears_proj <- qext$nYears_proj
proj_years  <- qext$proj_years
N_hist_q    <- qext$N_hist_q    # named list of nPop × nYears matrices (analytical delta-method)
r1          <- qext$qext_results[["thr20"]]   # threshold = 20 adults

# ── Per-population historical bands (simulation-based from N_hist_q) ─────────
df_hist_pop <- expand.grid(pop = seq_len(nPop),
                           y_idx = seq_len(nYears)) |>
  as_tibble() |>
  mutate(
    pop_name  = U_LocalPops_In[pop],
    core_name = U_CoreAreas_In[pop_core_vec[pop]],
    year      = DatYears[y_idx],
    N_p50     = N_hist_q$N_p50[cbind(pop, y_idx)],
    N_p05     = N_hist_q$N_p05[cbind(pop, y_idx)],
    N_p25     = N_hist_q$N_p25[cbind(pop, y_idx)],
    N_p75     = N_hist_q$N_p75[cbind(pop, y_idx)],
    N_p95     = N_hist_q$N_p95[cbind(pop, y_idx)]
  )

# ── Per-population projection bands ──────────────────────────────────────────
df_proj_pop <- expand.grid(pop = seq_len(nPop),
                           h   = seq_len(nYears_proj)) |>
  as_tibble() |>
  mutate(
    pop_name  = U_LocalPops_In[pop],
    core_name = U_CoreAreas_In[pop_core_vec[pop]],
    year      = proj_years[h],
    N_p05     = r1$N_p05[cbind(pop, h)],
    N_p25     = r1$N_p25[cbind(pop, h)],
    N_p50     = r1$N_p50[cbind(pop, h)],
    N_p75     = r1$N_p75[cbind(pop, h)],
    N_p95     = r1$N_p95[cbind(pop, h)]
  )

# ── Core-level projection: sum populations within each core ──────────────────
N_proj_arr <- exp(qext$logN_proj)   # nsim × nPop × nYears_proj
nSim       <- dim(N_proj_arr)[1]

df_proj_core <- bind_rows(lapply(seq_len(nCore), function(cc) {
  pops   <- which(pop_core_vec == cc)
  Nc_sim <- if (length(pops) == 1)
               matrix(N_proj_arr[, pops, ], nrow = nSim)
             else
               apply(N_proj_arr[, pops, ], c(1, 3), sum)
  Nq <- apply(Nc_sim, 2, quantile,
               probs = c(0.05, 0.25, 0.5, 0.75, 0.95), na.rm = TRUE)
  data.frame(core_name = U_CoreAreas_In[cc], year = proj_years,
             N_p05 = Nq[1,], N_p25 = Nq[2,], N_p50 = Nq[3,],
             N_p75 = Nq[4,], N_p95 = Nq[5,])
}))

# ── Core-level historical MAP abundance ──────────────────────────────────────
dfN_core <- expand.grid(pop = seq_len(nPop), y_idx = seq_len(nYears)) |>
  as_tibble() |>
  mutate(
    core_name   = U_CoreAreas_In[pop_core_vec[pop]],
    year_actual = DatYears[y_idx],
    N           = exp(logN_mat[cbind(pop, y_idx)])
  ) |>
  group_by(core_name, year_actual) |>
  summarise(N_total = sum(N, na.rm = TRUE), .groups = "drop")
```

```{r helper-functions, include=FALSE}
# ── Population trajectory plot (historical + projected) ──────────────────────
plot_trajectory <- function(pop_idx) {

  # Historical: delta-method CI on MAP logN
  se <- logN_se_mat[pop_idx, ]
  lN <- logN_mat[pop_idx, ]
  traj <- tibble(
    year   = DatYears,
    N      = exp(lN),
    N_lo90 = exp(lN - Z90 * se),
    N_hi90 = exp(lN + Z90 * se),
    N_lo50 = exp(lN - Z50 * se),
    N_hi50 = exp(lN + Z50 * se)
  )

  # Scale observed redd totals to adult-equivalent
  sites_p  <- which(site_pop_1idx == pop_idx)
  pS_total <- sum(data_list$pS[sites_p])
  obs <- obs_pop_year |>
    filter(pop == pop_idx) |>
    mutate(N_obs_equiv = obs_total / (0.5 * pS_total))

  # Projection bands
  d_proj <- df_proj_pop |> filter(pop == pop_idx)

  y_max <- max(
    max(traj$N_hi90,     na.rm = TRUE),
    max(obs$N_obs_equiv, na.rm = TRUE),
    max(d_proj$N_p95,    na.rm = TRUE),
    THRESHOLD_ADULTS * 1.5
  ) * 1.08

  ggplot() +
    # Historical CI ribbons
    geom_ribbon(data = traj,
                aes(x = year, ymin = N_lo90, ymax = N_hi90),
                fill = "steelblue", alpha = 0.15) +
    geom_ribbon(data = traj,
                aes(x = year, ymin = N_lo50, ymax = N_hi50),
                fill = "steelblue", alpha = 0.30) +
    geom_line(data = traj, aes(x = year, y = N),
              color = "steelblue4", linewidth = 1.1) +
    # Observed (redd-equivalent)
    geom_point(data = obs, aes(x = year, y = N_obs_equiv),
               shape = 21, fill = "white", color = "grey30",
               size = 2.2, stroke = 0.8, na.rm = TRUE) +
    # Divider
    geom_vline(xintercept = DatYears[nYears] + 0.5,
               linetype = 2, color = "grey50", linewidth = 0.4) +
    # Projection ribbons
    geom_ribbon(data = d_proj,
                aes(x = year, ymin = N_p05, ymax = N_p95),
                fill = "tomato", alpha = 0.15) +
    geom_ribbon(data = d_proj,
                aes(x = year, ymin = N_p25, ymax = N_p75),
                fill = "tomato", alpha = 0.30) +
    geom_line(data = d_proj, aes(x = year, y = N_p50),
              color = "tomato3", linewidth = 1.0) +
    # Management threshold
    geom_hline(yintercept = THRESHOLD_ADULTS,
               linetype = 3, color = "firebrick", linewidth = 0.8) +
    annotate("text",
             x     = min(DatYears) + 0.5,
             y     = THRESHOLD_ADULTS * 1.10,
             label = paste0(THRESHOLD_ADULTS, "-adult threshold"),
             color = "firebrick", size = 3.0, hjust = 0) +
    scale_y_continuous(labels = comma, limits = c(0, y_max)) +
    labs(
      x        = NULL,
      y        = "Adults (estimated / observed equivalent)",
      title    = "Estimated adult abundance",
      subtitle = paste0(
        "Historical: MAP ± 50/90% CI (blue)  |  Points: redds ÷ (0.5 × survey coverage)\n",
        "Projected ", nYears_proj, " yr: median + 50/90% CI (red)  |  ",
        "Future covariates: ", qext$future_cov_assumption
      )
    ) +
    theme_out
}

# ── Sensitivity forest plot ───────────────────────────────────────────────────
plot_sensitivity <- function(pops_in_core) {
  covars <- list(
    list(vals = bFlow_pop_vals, ses = bFlow_pop_se,
         mean = bFlow_M_val,   label = "Summer flow"),
    list(vals = bTemp_pop_vals, ses = bTemp_pop_se,
         mean = bTemp_M_val,   label = "August temperature"),
    list(vals = bINV_pop_vals,  ses = bINV_pop_se,
         mean = bINV_M_val,    label = "Non-native trout")
  )
  sens_df <- bind_rows(lapply(covars, function(cv) {
    tibble(
      pop_name   = pop_display[pops_in_core],
      covariate  = cv$label,
      estimate   = cv$vals[pops_in_core],
      se         = cv$ses[pops_in_core],
      hyper_mean = cv$mean
    )
  })) |>
    mutate(
      lo95      = estimate - 1.96 * se,
      hi95      = estimate + 1.96 * se,
      sig       = (lo95 > 0) | (hi95 < 0),
      pop_name  = factor(pop_name, levels = rev(pop_display[pops_in_core])),
      covariate = factor(covariate,
                         levels = c("Summer flow", "August temperature",
                                    "Non-native trout"))
    )

  ggplot(sens_df, aes(x = estimate, y = pop_name)) +
    geom_vline(xintercept = 0, linetype = 2, color = "grey50", linewidth = 0.7) +
    geom_vline(aes(xintercept = hyper_mean),
               linetype = 3, color = "goldenrod3", linewidth = 0.9) +
    geom_errorbarh(aes(xmin = lo95, xmax = hi95),
                   height = 0.3, alpha = 0.55, color = "grey40") +
    geom_point(aes(fill = sig), shape = 21, size = 2.8,
               color = "grey30", stroke = 0.5) +
    scale_fill_manual(values = c(`FALSE` = "white", `TRUE` = "steelblue3"),
                      guide = "none") +
    facet_wrap(~covariate, scales = "free_x", nrow = 1) +
    labs(
      x        = "Effect on log-λ per SD covariate",
      y        = NULL,
      title    = "Population sensitivity to dynamic covariates",
      subtitle = "Error bars: 95% CI  |  Gold dotted: global hyper-mean  |  Filled: CI excludes zero"
    ) +
    theme_out +
    theme(strip.text      = element_text(size = 9, face = "bold"),
          panel.spacing.x = unit(12, "pt"))
}

# ── Exposure time series ──────────────────────────────────────────────────────
plot_exposure <- function(pops_in_core, ca) {
  yr         <- DatYears
  pop_labels <- setNames(pop_display[pops_in_core], as.character(pops_in_core))

  mat_to_long <- function(mat, pops) {
    bind_rows(lapply(pops, function(p) {
      tibble(year = yr, value = mat[p, ], pop_name = pop_display[p])
    })) |> filter(!is.na(value) & is.finite(value))
  }

  df_f <- mat_to_long(FlowMat, pops_in_core)
  df_t <- mat_to_long(TempMat, pops_in_core)
  df_i <- mat_to_long(INVMat,  pops_in_core)

  n_pops_c   <- length(pops_in_core)
  pal        <- scales::hue_pal()(n_pops_c)
  pop_names  <- pop_display[pops_in_core]
  color_scale <- scale_color_manual(values = setNames(pal, pop_names), name = NULL)

  p_f <- ggplot(df_f, aes(x = year, y = value, color = pop_name)) +
    geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
    geom_line(linewidth = 0.75, alpha = 0.85) +
    color_scale + guides(color = "none") +
    labs(x = NULL, y = "z-score", title = "Summer streamflow (z, lag 6yr)") +
    theme_out

  p_t <- ggplot(df_t, aes(x = year, y = value, color = pop_name)) +
    geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
    geom_line(linewidth = 0.75, alpha = 0.85) +
    color_scale + guides(color = "none") +
    labs(x = NULL, y = "z-score", title = "August temperature (z, lag 6yr)") +
    theme_out

  p_i <- ggplot(df_i, aes(x = year, y = value, color = pop_name)) +
    geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
    geom_line(linewidth = 0.75, alpha = 0.85) +
    color_scale +
    labs(x = NULL, y = "standardized index", title = "Non-native trout") +
    theme_out +
    theme(legend.position  = "bottom",
          legend.text      = element_text(size = 7),
          legend.key.width = unit(16, "pt"))

  has_lkt_ca <- isTRUE(data_list$has_LKT[ca] == 1L)
  lkt_vals   <- if (has_lkt_ca) LKTMat[ca, ] else rep(NA_real_, nYears)
  df_l <- tibble(year = yr, value = lkt_vals) |>
    filter(!is.na(value) & is.finite(value))

  p_l <- if (nrow(df_l) > 0) {
    ggplot(df_l, aes(x = year, y = value)) +
      geom_hline(yintercept = 0, linetype = 2, color = "grey70", linewidth = 0.5) +
      geom_area(fill = "purple4", alpha = 0.15) +
      geom_line(color = "purple4", linewidth = 0.9) +
      labs(x = NULL, y = "standardized index", title = "Lake trout (core-level)") +
      theme_out
  } else {
    ggplot() +
      annotate("text", x = 0.5, y = 0.5,
               label = "No lake trout data\nfor this core area",
               size = 4, color = "grey60") +
      labs(title = "Lake trout (core-level)") + theme_out
  }

  (p_f | p_t) / (p_i | p_l)
}
```

This page presents model results organized by spatial level — global summary,
core area trajectories, and local population detail. Use the table of contents
on the right to navigate directly to any core area or population. For model
structure and diagnostics, see the [Model](model.html) tab.

---

## Global Results

Summary of model fit for the full multi-population dataset. A convergence code
of 0 indicates the optimizer reached a local minimum; the maximum absolute
gradient over fixed-effect parameters is the primary numerical convergence
criterion (values below 0.01 are generally acceptable).

```{r global-fit, results='asis'}
max_grad <- max(abs(fit_rep$gradient.fixed))
n_obs_   <- sum(!is.na(data_list$ReddCounts))
n_par    <- length(optB$par)

tibble(
  Metric = c("Convergence code", "Objective (NLL)", "Max |gradient|",
             "Site-year observations", "Outer parameters"),
  Value  = c(
    ifelse(optB$convergence == 0, "0 (relative convergence)", optB$convergence),
    round(optB$objective, 2),
    formatC(max_grad, format = "e", digits = 2),
    n_obs_,
    n_par
  )
) |>
  kable(col.names = c("Metric", "Value"), align = c("l","r")) |>
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE, font_size = 12)
```

### Covariate Effects

Estimated effects of each environmental and habitat covariate on population
growth rate (log-λ). Positive values indicate a beneficial effect; negative
values indicate suppression. Dynamic covariates (flow, temperature, invasives,
lake trout) are estimated as random slopes that vary across populations around
a global hyper-mean shown here.

```{r covar-table}
get_est <- function(nm) {
  v <- fit_rep$value[names(fit_rep$value) == nm]
  s <- fit_rep$sd[names(fit_rep$value)    == nm]
  if (!length(v)) return(c(NA_real_, NA_real_))
  c(v[1], s[1])
}

covar_rows <- list(
  list(par = "bFlow_M",  label = "Summer flow (hyper-mean)",        unit = "log-λ / SD flow"),
  list(par = "bTemp_M",  label = "August temperature (hyper-mean)", unit = "log-λ / SD temp"),
  list(par = "bTemp2",   label = "Temperature² (quadratic)",        unit = "log-λ / SD temp²"),
  list(par = "bINV_M",   label = "Non-native trout (hyper-mean)",   unit = "log-λ / SD INV"),
  list(par = "bLKT_M",   label = "Lake trout (global scalar)",      unit = "log-λ / SD LKT"),
  list(par = "bAdf",     label = "Lake area (adfluvial access)",    unit = "log-λ / log(km²+1)"),
  list(par = "bUVB",     label = "Unconfined valley bottom",        unit = "log-λ / z-score"),
  list(par = "bFlood",   label = "W95 flood frequency",             unit = "log-λ / z-score"),
  list(par = "bRD",      label = "Road density",                    unit = "log-λ / z-score")
)

covar_df <- bind_rows(lapply(covar_rows, function(r) {
  es <- get_est(r$par)
  tibble(Parameter = r$label, Units = r$unit, Estimate = es[1], SE = es[2])
})) |>
  filter(!is.na(Estimate)) |>
  mutate(
    z   = Estimate / SE,
    p   = 2 * pnorm(-abs(z)),
    Sig = case_when(
      abs(z) >= 2.576 ~ "***",
      abs(z) >= 1.960 ~ "**",
      abs(z) >= 1.645 ~ "*",
      TRUE            ~ ""
    ),
    across(c(Estimate, SE, z), ~ round(., 3)),
    p = round(p, 4)
  )

covar_df |>
  kable(col.names = c("Covariate", "Units", "Estimate", "SE", "z", "p", ""),
        align = c("l","l","r","r","r","r","l")) |>
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE, font_size = 12) |>
  footnote(general = "*** p < 0.01  ** p < 0.05  * p < 0.10.  Dynamic covariates carry an effective 6-yr lag.")
```

### Thermal Performance Curve

Bull trout population growth rate shows a unimodal (hump-shaped) response to
August water temperature: growth is fastest at intermediate temperatures and
declines at both extremes. The curve is estimated jointly from the linear and
quadratic temperature coefficients in the model.

```{r thermal-curve, fig.height=4}
if (covar_config$use_Temp == 1) {
  plot_temp_curve(fit_rep, temp_mean = temp_mean, temp_sd = temp_sd)
}
```

### Variance Components

How much do populations differ from one another, and how much year-to-year
noise is in the data? This table summarizes the estimated standard deviations
(SDs) of process noise, autocorrelation in annual deviations, observation
overdispersion, and the hierarchical random effects that allow demographic
parameters to vary across populations and core areas.

```{r variance-table}
var_rows <- list(
  list(par = "sigmaN",          label = "Process noise (σ_N)",               type = "Process"),
  list(par = "rho",             label = "AR(1) autocorrelation (ρ)",          type = "Process"),
  list(par = "phi_obs",         label = "NB2 dispersion (φ_obs)",             type = "Observation"),
  list(par = "sigma_b0r_pop",   label = "Growth rate SD — population",        type = "Hierarchy"),
  list(par = "sigma_b0r_core",  label = "Growth rate SD — core area",         type = "Hierarchy"),
  list(par = "sigma_b0phi_pop", label = "Density dependence SD — population", type = "Hierarchy"),
  list(par = "sigma_bFlow",     label = "Flow slope SD — population",         type = "Hierarchy"),
  list(par = "sigma_bTemp",     label = "Temp slope SD — population",         type = "Hierarchy"),
  list(par = "sigma_bINV",      label = "Invasives slope SD — population",    type = "Hierarchy")
)

var_df <- bind_rows(lapply(var_rows, function(r) {
  es <- get_est(r$par)
  tibble(Type = r$type, Parameter = r$label,
         Estimate = round(es[1], 3), SE = round(es[2], 3))
})) |>
  filter(!is.na(Estimate))

var_df |>
  kable(col.names = c("Type", "Parameter", "Estimate", "SE"),
        align = c("l","l","r","r")) |>
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE, font_size = 12) |>
  column_spec(1, bold = TRUE)
```

---

## Core-Area Abundance Trajectories

Historical estimated abundance (blue) and `r nYears_proj`-year stochastic
forward projection (red, 50% and 90% CI from `r qext$nsim` simulations).
Future covariates held at historical mean (z-score = 0).

```{r core-trajectories, fig.height=14}
ggplot() +
  geom_area(data = dfN_core,
            aes(x = year_actual, y = N_total),
            fill = "steelblue", alpha = 0.15) +
  geom_line(data = dfN_core,
            aes(x = year_actual, y = N_total),
            color = "steelblue", linewidth = 0.9) +
  geom_vline(xintercept = DatYears[nYears] + 0.5,
             linetype = 2, color = "grey50", linewidth = 0.4) +
  geom_ribbon(data = df_proj_core,
              aes(x = year, ymin = N_p05, ymax = N_p95),
              fill = "tomato", alpha = 0.15) +
  geom_ribbon(data = df_proj_core,
              aes(x = year, ymin = N_p25, ymax = N_p75),
              fill = "tomato", alpha = 0.30) +
  geom_line(data = df_proj_core,
            aes(x = year, y = N_p50),
            color = "tomato", linewidth = 0.9) +
  facet_wrap(~core_name, scales = "free_y", ncol = 4) +
  labs(x = "Year", y = "Estimated spawners (sum of local populations)",
       title = "Estimated Bull Trout Spawner Abundance by Core Area",
       subtitle = paste0(
         "Historical: MAP estimate (blue)  |  Projected ", nYears_proj,
         " years: median + 50/90% CI (red)  |  Future covariates: ",
         qext$future_cov_assumption
       )) +
  theme_bw(base_size = 11) +
  theme(strip.text = element_text(size = 8),
        panel.grid.minor = element_blank())
```

---

## Demographic Risk Summary

Probability that a local population falls below 20 adults at any point during
the `r nYears_proj`-year forward projection, based on `r qext$nsim` stochastic
simulations. Populations near or below this threshold are at elevated risk of
local extirpation. Results are sorted from highest to lowest risk.

```{r qext-table}
qext_tbl20 <- qext$qext_table |>
  filter(threshold == 20) |>
  arrange(desc(p_qext)) |>
  select(pop_name, core_name, p_qext, N_med_end, N_lo_end, N_hi_end) |>
  mutate(pop_name = pop_display[match(pop_name, U_LocalPops_In)])

qext_tbl20 |>
  kable(col.names = c("Population", "Core Area",
                       "P(N < 20)", "N median", "N 5%", "N 95%"),
        align = c("l","l","r","r","r","r"),
        digits = c(0, 0, 3, 1, 1, 1)) |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                font_size = 12, full_width = FALSE) |>
  column_spec(3, bold = TRUE,
              color = ifelse(qext_tbl20$p_qext > 0.5, "firebrick", "black"))
```

---

## Population Trajectories

Each section below shows the core-level growth rate, population-level sensitivity
to dynamic covariates, covariate exposure time series, and individual population
abundance trajectories (historical MAP ± CI + `r nYears_proj`-year projection).

```{r core-index, results='asis'}
cat("### Jump to core area\n\n")
for (ca in seq_len(nCore)) {
  n_pops <- sum(pop_core_vec == ca)
  cat(sprintf("- [%s](#%s) — %d population%s\n",
              U_CoreAreas_In[ca], core_anchor[ca],
              n_pops, if (n_pops > 1) "s" else ""))
}
cat("\n---\n\n")
```

```{r all-sections, results='asis', fig.width=10, fig.height=7}
b0r_core_vals <- fit_rep$value[names(fit_rep$value) == "b0r_core"]
b0r_core_se   <- fit_rep$sd[names(fit_rep$value)    == "b0r_core"]
b0r_pop_vals  <- fit_rep$value[names(fit_rep$value) == "b0r_pop"]
b0r_pop_se    <- fit_rep$sd[names(fit_rep$value)    == "b0r_pop"]

for (ca in seq_len(nCore)) {

  core_label   <- U_CoreAreas_In[ca]
  pops_in_core <- which(pop_core_vec == ca)
  n_pops       <- length(pops_in_core)

  cat(sprintf("\n\n## %s {#%s}\n\n", core_label, core_anchor[ca]))

  yr_range_core <- range(DatYears[apply(
    data_list$ReddCounts[which(site_pop_1idx %in% pops_in_core), , drop = FALSE],
    2, function(x) any(!is.na(x))
  ) |> which()])

  cat(sprintf("**%d local population%s** | Data span: **%d – %d**\n\n",
              n_pops, if (n_pops > 1) "s" else "",
              yr_range_core[1], yr_range_core[2]))

  core_b0r    <- b0r_core_vals[ca]
  core_b0r_se <- b0r_core_se[ca]
  if (!is.na(core_b0r)) {
    cat(sprintf(
      "> **Core growth rate** (b0r_core): **%.3f** (SE: %.3f) — λ ≈ **%.3f**\n\n",
      core_b0r, core_b0r_se, exp(core_b0r)
    ))
  }

  # Population growth rate table
  pop_summary <- tibble(
    Population    = pop_display[pops_in_core],
    `b0r (pop)`   = round(b0r_pop_vals[pops_in_core], 3),
    `SE`          = round(b0r_pop_se[pops_in_core],   3),
    `λ (exp b0r)` = round(exp(b0r_pop_vals[pops_in_core]), 3)
  )
  print(
    kable(pop_summary, align = c("l","r","r","r")) |>
    kable_styling(bootstrap_options = c("striped","condensed"),
                  font_size = 11, full_width = FALSE)
  )
  cat("\n\n")

  cat("#### Covariate sensitivity\n\n")
  print(plot_sensitivity(pops_in_core))
  cat("\n\n")

  cat("#### Covariate exposure\n\n")
  print(plot_exposure(pops_in_core, ca))
  cat("\n\n<hr>\n\n")

  for (pop_idx in pops_in_core) {
    cat(sprintf("\n\n### %s {#%s}\n\n",
                pop_display[pop_idx], pop_anchor[pop_idx]))
    print(plot_trajectory(pop_idx))
    cat("\n\n")
  }
}
```
